---
pagetitle: "NTEE Codes with NAICS Crosswalk"
format:
  html:
    page-layout: full
    css: ntee-table.css
    theme: none
    minimal: true
    embed-resources: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Data source: the contracted NCCS lookup tables on S3 (nccs-contracts ADR 0049).
# Set NTEE_LOOKUP_DIR to a local folder holding the two CSVs to render before
# a lookup publish has landed; otherwise the live files are read.

lookup_base_url <- "https://nccsdata.s3.amazonaws.com/lookups/bmf/latest"
lookup_dir      <- Sys.getenv("NTEE_LOOKUP_DIR", unset = "")

lookup_source <- function(file_name) {
  if (nzchar(lookup_dir)) file.path(lookup_dir, file_name) else paste(lookup_base_url, file_name, sep = "/")
}

download_url <- paste(lookup_base_url, "ntee_code.csv", sep = "/")

read_lookup_csv <- function(file_name) {
  readr::read_csv(
    lookup_source(file_name),
    col_types = readr::cols(.default = readr::col_character())
  )
}

ntee_codes   <- read_lookup_csv("ntee_code.csv")
major_groups <- read_lookup_csv("ntee_code_major_group.csv")

vintage_label <- tryCatch({
  manifest <- jsonlite::fromJSON(lookup_source("_manifest.json"))
  paste0(manifest$vintage, ", built ", substr(manifest$built_at, 1, 10))
}, error = function(e) "current vintage")

crosswalk <- ntee_codes |>
  dplyr::mutate(major_group_letter = substr(ntee_code, 1, 1)) |>
  dplyr::left_join(
    major_groups |>
      dplyr::select(major_group_letter = ntee_code_first_character, ntee_code_major_group),
    by = "major_group_letter"
  ) |>
  dplyr::transmute(
    Code          = ntee_code,
    `Major group` = factor(paste0(major_group_letter, ": ", dplyr::coalesce(ntee_code_major_group, "Unknown"))),
    Description   = ntee_code_definition,
    `NAICS code`  = dplyr::if_else(naics_code == "UNDEFINED", "", naics_code),
    Details       = dplyr::coalesce(ntee_code_description, "")
  ) |>
  dplyr::arrange(Code)

n_codes <- nrow(crosswalk)
```

::: {.ntee-page}

# NTEE Codes with NAICS Crosswalk

<p class="lede">Every National Taxonomy of Exempt Entities (NTEE) code the IRS assigns, with its short and long description and the closest North American Industry Classification System (NAICS, 2022 edition) industry. This is the same table the NCCS data pipeline uses, so it matches the codes in the Business Master File. Codes the IRS has retired are kept because older files still carry them.</p>

<div class="ntee-actions">
  <a class="ntee-button -primary" href="`r download_url`" download>Download the full table (CSV)</a>
  <a class="ntee-button" href="../../datasets/ntee/">About this data set</a>
  <a class="ntee-button" href="../../resources/ntee/">NTEE guide</a>
  <span class="ntee-meta">`r n_codes` codes &middot; vintage `r vintage_label`</span>
</div>

```{r table}
DT::datatable(
  crosswalk,
  rownames   = FALSE,
  filter     = "top",
  escape     = TRUE,
  extensions = "Buttons",
  plugins    = "ellipsis",
  options    = list(
    pageLength = 25,
    lengthMenu = c(25, 50, 100, 700),
    dom        = "Bfrtip",
    buttons    = list(
      list(extend = "csv",  text = "Download shown rows (CSV)",  filename = "ntee-naics-crosswalk-selection"),
      list(extend = "copy", text = "Copy")
    ),
    columnDefs = list(
      list(targets = 0, className = "ntee-code"),
      list(targets = 4, className = "ntee-long",
           render = DT::JS("$.fn.dataTable.render.ellipsis(160, false)"))
    ),
    language = list(search = "Search all columns:")
  )
)
```

<p class="ntee-meta">Source: NCCS lookup tables, <code>lookups/bmf/latest/ntee_code.csv</code>. Short descriptions are the IRS wording (Instructions for Form 1023, Appendix D). Long descriptions come from the Nonprofit Open Data Collective NTEE table; the newest IRS codes have none yet. Hover over a shortened description to read all of it.</p>

:::
