---
title: "E-filed 990s are up 1.5x since 2015, and 60k+ new public charities have grown into the full Form 990"
date: 2026-07-24
description: "How many 990s were e-filed for tax year 2015 versus 2023, and how many public charities founded 2015-2023 have filed a full Form 990, from canonical IRS/NCCS data."
format: gfm
type: analysis
categories:
  - efile
  - trends
author:
  - id: thiya
citation:
  container-title: National Center for Charitable Statistics
  volume: 1
  issue: 1
  note: >-
    E-file v2.1 990/990-EZ headers (efile_v2_1@2009-2024); BMF master
    (geocoded, rolling@2026-07-24). Full provenance in _pins.csv of
    nccs-data-requests/requests/2026-07-efile-growth-new-charities.
# No figures in this doc; png avoids the cairo/X11 dependency of the project
# default fig-format svg on machines without XQuartz.
knitr:
  opts_chunk:
    dev: png
---

```{r setup, include=FALSE}
library(dplyr)
source(here::here("R", "request_read.R"))
req <- here::here("requests", "2026-07-efile-growth-new-charities")
fmt <- function(x) format(x, big.mark = ",")
```

## The ask

> 1. How many 990s were e-filed in tax year 2015 versus tax year 2023?
> 2. How many new public charities, eligible for filing the full Form 990,
>    were founded between 2015 and 2023?

(Requester identity and the verbatim ask are in the gitignored `_private.md`.)

## Data & method

Two canonical artifacts, joined on `EIN2`:

- **E-file Form 990 header tables** (`s3://nccs-efile/public/efile_v2_1/`,
  `F9-P00-T00-HEADER-{year}.CSV`, all 16 published tax years 2009-2024).
  One row per e-filed 990 or 990-EZ. Two fields matter beyond identity:
  `RETURN_TYPE` distinguishes the full 990 from the 990-EZ, and
  `F9_00_YEAR_FORMATION` is the organization's founding year as
  self-reported on the return, a field the BMF does not carry (the BMF has
  only the IRS *ruling* date, which lags founding).
- **Geocoded unified BMF** (via `nccsdata::nccs_read`), which supplies
  public-charity status: `org_type = "public_charity"` is the package's
  canonical cut (501(c)(3), foundation code not private-foundation).

Conventions this analysis follows:

- **Everything is by tax year.** NCCS deliberately re-partitions IRS
  calendar-year files by tax year during harmonization; all counts below are
  tax-year counts.
- **Form type and tax year are taken from the `RETURN_TYPE` and `TAX_YEAR`
  columns on every row**, never inferred from a file's name. The read
  helper carries a `file_year` column so the render itself verifies the
  partitioning (see the check below).
- "Eligible for filing the full 990" is proxied by *demonstrated* filing:
  an EIN counts if it filed at least one full Form 990 (not EZ) in tax
  years 2015-2024. A threshold-based alternative (receipts >= $200k or
  assets >= $500k in the current BMF) is reported as a sensitivity check.

```{r read, include=FALSE}
hdr <- read_efile_header(
  req, years = 2009:2024,
  columns = c("EIN2", "RETURN_TYPE", "TAX_YEAR", "F9_00_YEAR_FORMATION"),
  note = "990/990-EZ headers, all published tax years"
)

pc <- read_bmf_master(
  req, org_type = "public_charity",
  columns = c("EIN2", "foundation_code", "ruling_date_ym_str",
              "income_amount", "asset_amount"),
  cache = FALSE,
  note = "public charities (canonical org_type cut), full registry"
)
```

```{r verify, include=FALSE}
# Column-verified partitioning: every row's TAX_YEAR must match its file.
n_mismatch <- sum(hdr$TAX_YEAR != hdr$file_year, na.rm = TRUE)
rt_values  <- sort(unique(hdr$RETURN_TYPE))
stopifnot(identical(rt_values, c("990", "990EZ")))
```

Partitioning check on this render: `r fmt(nrow(hdr))` header rows scanned,
**`r n_mismatch` rows** where the in-column `TAX_YEAR` disagrees with the
file-name year; `RETURN_TYPE` values present: `r paste(rt_values, collapse = ", ")`.

## Findings

### Q1: E-filed 990s by tax year

```{r q1}
q1 <- hdr |>
  count(TAX_YEAR, RETURN_TYPE) |>
  group_by(TAX_YEAR) |>
  summarise(
    `Full 990` = sum(n[RETURN_TYPE == "990"]),
    `990-EZ`   = sum(n[RETURN_TYPE == "990EZ"]),
    Total      = sum(n),
    .groups = "drop"
  ) |>
  arrange(TAX_YEAR)
knitr::kable(q1 |> mutate(across(-TAX_YEAR, fmt)), align = "lrrr")
```

```{r q1-inline, include=FALSE}
t15 <- q1$Total[q1$TAX_YEAR == 2015]
t23 <- q1$Total[q1$TAX_YEAR == 2023]
```

**Tax year 2015: `r fmt(t15)` e-filed returns. Tax year 2023: `r fmt(t23)`**,
about `r round(t23 / t15, 1)`x. Two structural notes: e-filing was optional
before tax year 2021 (the Taxpayer First Act mandate), so early years
reflect e-file adoption rather than all filings, while 2023 is
near-complete; and tax year 2024 is still accruing. This table covers the
990 and 990-EZ; the 990-PF and 990-N are separate products.

### Q2: New public charities that grew into the full 990

```{r q2}
founders <- hdr |>
  filter(
    RETURN_TYPE == "990",
    !is.na(F9_00_YEAR_FORMATION),
    F9_00_YEAR_FORMATION >= 2015, F9_00_YEAR_FORMATION <= 2023,
    F9_00_YEAR_FORMATION <= TAX_YEAR   # formation cannot postdate the tax year
  ) |>
  group_by(EIN2) |>
  summarise(formation_year = min(F9_00_YEAR_FORMATION), .groups = "drop")

new_pc <- founders |> semi_join(pc, by = "EIN2")

knitr::kable(
  new_pc |> count(formation_year, name = "Public charities") |>
    mutate(`Public charities` = fmt(`Public charities`)),
  align = "lr"
)
```

```{r q2-sensitivity, include=FALSE}
# Strict 170(b)(1)(A) / 509(a) foundation-code cut (excludes missing/suspense codes)
strict_codes <- c("10","11","12","13","14","15","16","17","18",
                  "21","22","23","24","25")
n_strict <- new_pc |> semi_join(pc |> filter(foundation_code %in% strict_codes),
                                by = "EIN2") |> nrow()

# BMF-only alternative: ruling year + statutory full-990 thresholds
pc_ruled <- pc |>
  mutate(ruling_year = as.integer(substr(ruling_date_ym_str, 1, 4))) |>
  filter(ruling_year >= 2015, ruling_year <= 2023)
n_ruled  <- nrow(pc_ruled)
n_thresh <- pc_ruled |>
  filter(suppressWarnings(as.numeric(income_amount)) >= 2e5 |
         suppressWarnings(as.numeric(asset_amount))  >= 5e5) |>
  nrow()
```

**`r fmt(nrow(new_pc))` public charities** founded 2015-2023 have filed at
least one full Form 990 (of `r fmt(nrow(founders))` full-990 filers
reporting those formation years across all subsections).

Sensitivity and cross-checks:

- Restricting to explicit 170(b)(1)(A) / 509(a) foundation codes (dropping
  orgs whose foundation code is missing or in suspense) gives
  `r fmt(n_strict)`.
- The BMF-only alternative: `r fmt(n_ruled)` public charities have IRS
  *ruling* years 2015-2023, of which `r fmt(n_thresh)` currently meet the
  statutory full-990 thresholds (receipts >= $200k or assets >= $500k).
  Two independent definitions land in the same range, which is reassuring.

### How much of the e-file growth is new organizations?

The anticipated follow-up: did full-990 e-filers grow because new charities
formed, or because existing charities were not e-filing in 2015?

```{r q3}
new_by_ty <- hdr |>
  filter(
    RETURN_TYPE == "990",
    !is.na(F9_00_YEAR_FORMATION),
    F9_00_YEAR_FORMATION >= 2015, F9_00_YEAR_FORMATION <= 2023,
    F9_00_YEAR_FORMATION <= TAX_YEAR
  ) |>
  distinct(TAX_YEAR, EIN2)

decomp <- hdr |>
  filter(RETURN_TYPE == "990") |>
  count(TAX_YEAR, name = "full_990_filers") |>
  left_join(
    new_by_ty |> count(TAX_YEAR, name = "founded_2015_plus"),
    by = "TAX_YEAR"
  ) |>
  left_join(
    new_by_ty |> semi_join(pc, by = "EIN2") |>
      count(TAX_YEAR, name = "of_which_public_charities"),
    by = "TAX_YEAR"
  ) |>
  mutate(founded_pre_2015 = full_990_filers - founded_2015_plus) |>
  filter(TAX_YEAR >= 2015, TAX_YEAR <= 2023)

knitr::kable(decomp |> mutate(across(-TAX_YEAR, fmt)), align = "lrrrr")
```

```{r q3-inline, include=FALSE}
n15_new <- decomp$founded_2015_plus[decomp$TAX_YEAR == 2015]
n23_new <- decomp$founded_2015_plus[decomp$TAX_YEAR == 2023]
n15_old <- decomp$founded_pre_2015[decomp$TAX_YEAR == 2015]
n23_old <- decomp$founded_pre_2015[decomp$TAX_YEAR == 2023]
```

Between TY2015 and TY2023 the annual count of full-990 e-filers rose by
`r fmt((n23_new + n23_old) - (n15_new + n15_old))`. Organizations founded in
2015 or later account for `r fmt(n23_new - n15_new)` of that increase
(`r fmt(n23_new)` of TY2023's filers, versus `r fmt(n15_new)` in TY2015);
organizations founded before 2015 account for the remaining
`r fmt(n23_old - n15_old)`. Roughly half and half.

The coverage context reframes both halves: per the published e-file catalog,
only about **68% of 990/990-EZ filings were electronic in tax year 2015**,
rising to essentially 100% by 2023 (see the paper-vs-electronic table in
[catalog-efile-v2_1](https://urbaninstitute.github.io/nccs/catalogs/catalog-efile-v2_1.html)).
Adjusting TY2015's 358,413 e-filed returns for the missing paper share
implies a total filing volume close to TY2023's 527,764. Total 990/990-EZ
filing volume was therefore roughly flat: the pre-2015 organizations in the
table above are mostly *newly visible* rather than newly filing, and the
new organizations are largely replacing organizations that dissolved.

## Takeaways

- E-filed 990/990-EZ volume grew ~1.5x from tax year 2015 to 2023, with the
  mandate visible as a step in 2020-2021. Coverage-adjusted (68% electronic
  in 2015 vs ~100% in 2023), total filing volume was roughly flat: the
  growth in *visible* filers is predominantly the e-file mandate.
- Roughly 60-62k public charities founded since 2015 have already grown
  into the full Form 990. The count tapers for recent cohorts (2022-2023)
  purely from filing lag: young orgs have not yet filed their first 990.
- Of the growth in annual full-990 e-filers, about half is organizations
  founded 2015+ entering the data and half is pre-2015 organizations
  becoming newly visible; with flat total volume, new entrants largely
  replace dissolved organizations rather than expanding the total.
- Formation year is self-reported on the return; rows where it postdates
  the tax year are dropped as misreports.
- Orgs that only ever paper-filed (possible before tax year 2021) or only
  filed the EZ/N despite full-990 size are not counted, so the e-file-based
  number is a floor.

---

*Reproducibility: this story reads the vintages pinned in `_pins.csv`; rerun
from those pins to reproduce. First render downloads ~3.7 GB of e-file
headers into the gitignored `data/` cache; later renders are fast. See ADR
0024 / 0025 in `nccs-contracts`.*
