---
title: "nccsdata Part 4: Summary Tables"
date: 2023-11-03
description: Part 4 of 4 data stories covering the nccsdata R package. This story focuses on summarising NCCS legacy data.
featured: false
primaryCtaUrl: https://urbaninstitute.github.io/nccsdata
primaryCtaText: Package Website
format: gfm
type: methods
categories:
  - R packages
author:
  - id: thiya
citation: 
  container-title: National Center for Charitable Statistics
  volume: 1
  issue: 1
links:
  - header: Data Stories in this series
    links:
    - text: "Part 1: Downloading Data"
      href: https://urbaninstitute.github.io/nccs/stories/nccsdata/
      icon: article
    - text: "Part 2: NTEE Codes"
      href: https://urbaninstitute.github.io/nccs/stories/nccsdata-ntee/
      icon: article
    - text: "Part 3: Geographic Filters"
      href: https://urbaninstitute.github.io/nccs/stories/nccsdata-geo/
      icon: article
    - text: "Part 4: Summarising Data"
      href: https://urbaninstitute.github.io/nccs/stories/nccsdata-summary/
      icon: article
  - header: Package Links
    links:
    - text: Github Repository
      href: https://github.com/UrbanInstitute/nccsdata
      icon: github
    - text: Package Reference
      href: https://urbaninstitute.github.io/nccsdata/reference/index.html
      icon: link
    - text: Relevant Vignette
      href: https://urbaninstitute.github.io/nccsdata/articles/summary_stats.html
      icon: article
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
library(nccsdata)
library(dplyr)
```

## Introduction

In the final segment of our four-part series on the [`nccsdata`](https://urbaninstitute.github.io/nccsdata/) package, we construct tables summarizing NCCS Legacy Data.

## Create Summary Tables

After preparing data using [`get_data()`](https://urbaninstitute.github.io/nccsdata/reference/get_data.html) and subsequent processing through additional filtering or wrangling, you can use [`preview_sample()`](https://urbaninstitute.github.io/nccsdata/reference/preview_sample.html) to compute summary statistics for legacy datasets.

The following example illustrates this process:

```{r warning = FALSE}
core <- get_data(dsname = "core",
                 time = "2015")
```

```{r warning = FALSE}
preview_sample(data = core,
              group_by = c("NTEECC", "STATE"),
              var = c("TOTREV"),
              stats = c("count", "mean", "max"))
```



[`preview_sample()`](https://urbaninstitute.github.io/nccsdata/reference/preview_sample.html) groups the dataset based on the columns specified in `group_by`, and computes summary statistics for the user-defined *var* column. The available summary statistics are:

 * `min`: minimum value
 * `median`: median value
 * `max`: maximum value
 * `mean`: mean value
 * `count`: count of rows belonging to group

As long as the columns in *group_by* are present in the data set, [`preview_sample()`](https://urbaninstitute.github.io/nccsdata/reference/preview_sample.html) will be able to generate a summary table for the user.

## Conclusion

We hope this series has been helpful for working with the [`nccsdata`](https://urbaninstitute.github.io/nccsdata/index.html) package. package. Contact the package maintainers with any comments or suggestions and any additional features you would find useful.
