---
title: Payroll Taxes
date: 2023-08-28
description: This data story outlines the process for estimating the amount of payroll taxes nonprofits pay annually.
featured: true
featuredOrder: 1
type: methods
categories:
  - SOI extracts
  - payroll tax
author:
  - id: hmartin
  - id: jlecy
citation: 
  container-title: National Center for Charitable Statistics
  volume: 1
  issue: 1
  doi: 10.5555/12345678
links:
  - header: Replication Files
    links:  
    - text: Data + Code
      href: https://tinyurl.com/ylydvtj9
      icon: download
    - text: Code
      href: https://raw.githubusercontent.com/UrbanInstitute/nccs-recipes/main/vignettes/payroll/payroll.R
      icon: link
  - header: In Context 
    links:
    - text: Nonprofit Infrastructure Coalition 
      href: https://independentsector.org/wp-content/uploads/2023/04/nic-agenda-2023.pdf
      icon: download
---


The nonprofit sector must persistently track programs and activities in order to measure societal benefits and convey impact to key stakeholders. This research note introduces an uncommon but useful approach to measuring sector size though payroll tax information disclosed on IRS form 990s. 

Researchers commonly use counts of nonprofits or total program expenses to measure sector size. Alternatively payroll captures the mobilization of human resources, a complementary way to convey the depth of engagement within communities. 

Historically, payroll data has been a difficult to acquire. Partnerships between nonprofit scholars and the Bureau of Labor Statistics have resulted in new methodologies for using the Quarterly Census of Employment and Wages to measure the size of the nonprofit workforce, but QCEW data on nonprofits is currently only updated once every five years. Payroll tax information on 990 forms is more accessible and timely.

Although nonprofits are exempt from many types of taxes they are still required to pay federal payroll taxes. These include a 6.2% Social Security tax and 1.45% for Medicare. Taxes are levied for both part and full-time employees. The aggregate level of payroll taxes paid by the sector can be used as a proxy for the number of full-time equivalent workers employed by the sector.  

We estimate that nonprofits currently pay approximately $66 billion per year in payroll taxes. See below for details about how we arrive at this estimate. 

### Form 990 Data

We estimate the payroll taxes that the nonprofit sector contributes annually using 2019 IRS 990 data, the most recent full year of available data at the time the analysis was conducted. 

Three years of IRS 990 data extracts must be utilized to capture a single full year of data since there can be a 2-year lag in returns, resulting in an idiosyncratic data aggregation challenge. Form 990 data is compiled from the 2019, 2020, and 2021 Statistics of Income Extract files available through the IRS:
<https://www.irs.gov/statistics/soi-tax-stats-annual-extract-of-tax-exempt-organization-financial-data>. 


### Sampling Considerations

Data availability varies greatly by the size of the nonprofit. Larger nonprofits are required to file the full Form 990, which requires significant disclosures on programs and finances. Medium-sized nonprofits file a more compact version of the form called the 990EZ, and small nonprofits are only required to report that they remain active by filing Form 990N. Private foundations utilize the 990-PF. Payroll information disclosures vary by the type of form. 

***Larger Nonprofits:*** The Form 990 collects data on the amount of payroll taxes that larger nonprofits (i.e., those with gross receipts >= $200K or total assets >= $500K) contribute (**$55B in 2019**).

***Private foundations:*** The Form 990-PF does not collect data on the amount of payroll taxes that private foundations contribute However, it does collect data on the amount that private foundations spend on salaries and benefits ($3B). The Form 990 collects this data for larger nonprofits, as well. Therefore, we can divide the amount of payroll taxes that larger nonprofits contribute by the amount that larger nonprofits spend on payroll taxes to calculate a payroll tax/salaries and benefits ratio (5.75%). We can then apply that ratio to the salaries and benefits data from the Form 990-PF to estimate the amount of payroll taxes that private foundations contribute.

However, the IRS's Form 990-PF data extract only includes the columns with data on "compensation of officers, directors, trustees, etc." and "pension plans, employee benefits;" it excludes the column with data on "other employee salaries and wages." Therefore, we will underestimate the amount of payroll taxes that private foundations contribute. Additionally, the IRS does not provide a data extract for the 2019 Form 990-PF file. This creates a further underestimation of the amount of payroll taxes that private foundations contribute (**estimated $153M in 2019**).

***Smaller Nonprofits:*** The Form 990-EZ does not collect data on the amount of payroll taxes that smaller nonprofits (i.e., those with gross receipts <$200K and total assets <$500K) contribute It does collect data on the amount that smaller nonprofits spend on salaries and benefits, but the IRS's Form 990-EZ data extracts exclude these data. Therefore, we are unable to estimate the amount of payroll taxes that smaller nonprofits contribute.

***Invisible Nonprofits:*** Some nonprofits, such as churches, do not have to file annual returns. See <https://www.irs.gov/charities-non-profits/annual-exempt-organization-return-who-must-file>.

The Form 990-N also does not collect any financial data from the smallest nonprofits (i.e., those with gross receipts <=$50K), nor does it collect data on the amount that these nonprofits spend on salaries and benefits. Therefore, we are unable to estimate the amount of payroll taxes that the smallest nonprofits contribute.


## Data Vignette 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
```


### Packages 

You can install R packages as follows:

```{r, eval=F}
install.packages( "knitr" )
install.packages( "tidyverse" )
install.packages( "pander" )
install.packages( "kableExtra" )
```

If you already have these packages, you only need to load the libraries.

```{r}
library( tidyverse )  # data wrangling
library( pander )     # pretty formats
library( knitr )      # pretty formats
library( kableExtra ) # html formats for tables 
```


### Data 

Load the SOI Extract data obtained from the IRS: 

```{r, echo=F, eval=T, cache=T}
url.pc.2019 <- "C:/Users/jdlec/Downloads/payroll/19eoextract990.csv"
url.pc.2020 <- "C:/Users/jdlec/Downloads/payroll/20eoextract990.csv"
url.pc.2021 <- "C:/Users/jdlec/Downloads/payroll/21eoextract990.csv"
url.pf.2020 <- "C:/Users/jdlec/Downloads/payroll/20eoextract990pf.csv"
url.pf.2021 <- "C:/Users/jdlec/Downloads/payroll/21eoextract990pf.csv"

pc.2019 <- readr::read_csv( url.pc.2019 )
pc.2020 <- readr::read_csv( url.pc.2020 )
pc.2021 <- readr::read_csv( url.pc.2021 )
pf.2020 <- readr::read_csv( url.pf.2020 )
pf.2021 <- readr::read_csv( url.pf.2021 )
```

Download all data and read locally: 

```{r, eval=F}
URL <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/payroll.zip"
download.file( URL, destfile="payroll.zip" )
unzip( "payroll.zip" )

pc.2019 <- readr::read_csv( "payroll/19eoextract990.csv" )
pc.2020 <- readr::read_csv( "payroll/20eoextract990.csv" )
pc.2021 <- readr::read_csv( "payroll/21eoextract990.csv" )
pf.2020 <- readr::read_csv( "payroll/20eoextract990pf.csv" )
pf.2021 <- readr::read_csv( "payroll/21eoextract990pf.csv" )
```

Or read directly from the URLs: 

```{r, eval=F}
url.pc.2019 <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/19eoextract990.csv"
url.pc.2020 <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/20eoextract990.csv"
url.pc.2021 <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/21eoextract990.csv"
url.pf.2020 <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/20eoextract990pf.csv"
url.pf.2021 <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/21eoextract990pf.csv"

pc.2019 <- readr::read_csv( url.pc.2019 )
pc.2020 <- readr::read_csv( url.pc.2020 )
pc.2021 <- readr::read_csv( url.pc.2021 )
pf.2020 <- readr::read_csv( url.pf.2020 )
pf.2021 <- readr::read_csv( url.pf.2021 )
```



Preview the data: 

```{r}
head( pc.2019[,1:6] ) %>% pander()
```


The data dictionaries are available in the zipped directory. Additional information can be obtained at the [SOI Extracts page](https://www.irs.gov/statistics/soi-tax-stats-annual-extract-of-tax-exempt-organization-financial-data).


## Data Preparation 

### Variable Cleanup

Convert six digit tax period dates (YYYYMM) to four digit years (YYYY): 




```{r}
pc.2021$year <- pc.2021$tax_pd  %>% substr( 1, 4 )
pc.2020$year <- pc.2020$tax_pd  %>% substr( 1, 4 )
pc.2019$year <- pc.2019$tax_pd  %>% substr( 1, 4 )
pf.2020$year <- pf.2020$TAX_PRD %>% substr( 1, 4 )
pf.2021$year <- pf.2021$TAX_PRD %>% substr( 1, 4 )
```

```{r, results="hold"}
head(pc.2021$tax_pd)
head(pc.2021$year)
```

### Isolate the Study Period

SOI Extract files are organized by filing dates, the dates the returns were received by the IRS. The filing date is usually around six months after the end of an organization's fiscal year, which is also twelve months ahead of the calendar year that correspond to the nonprofits activities. 

In addition, nonprofits can back-file late returns or submit amended returns, so although the year of the SOI extract file is different from the periods of data contained within: 

```{r}
pc.2021 %>%
  group_by( year ) %>%
  summarise( n() ) %>%
  kable( col.names = c( "Tax Period", "Frequency" ),
         caption = "Tax Periods in the 2021 Form 990 Dataset", 
         align="c" ) 
```


```{r}
pf.2021 %>%
  group_by( year ) %>%
  summarise( n() ) %>%
  kable( col.names = c( "Tax Period", "Frequency" ),
         caption = "Tax Periods in the 2021 Form 990-PF Dataset",
         align="c" ) 
```


We need data that corresponds to the 2019 calendar year, the most recent full year available at the time of writing. 

```{r}
pc.2021.subset <- filter( pc.2021, year=="2019" )
pf.2021.subset <- filter( pf.2021, year=="2019" )
pc.2020.subset <- filter( pc.2020, year=="2019" )
pf.2020.subset <- filter( pf.2020, year=="2019" )
pc.2019.subset <- filter( pc.2019, year=="2019" )
```

Inspect to ensure we have the correct data now: 

```{r}
pc.2021.subset %>%
  group_by( year ) %>%
  summarise( n() ) %>%
  kable( col.names = c( "Tax Period", "Frequency" ), align="c",
         caption = "Tax Periods in the Subsetted 2021 Form 990 Dataset" )

pf.2021.subset %>%
  group_by( year ) %>%
  summarise( n() ) %>%
  kable( col.names = c( "Tax Period", "Frequency" ), align="c",
         caption = "Tax Periods in the Subsetted 2021 Form 990-PF Dataset" ) 
```


### Harmonizing the Data

In order to combine files we need to ensure that the variable names are the same. 

R is case-sensitive; if one dataset uses uppercase letters for a column name and another dataset uses lowercase letters for a column name, it will not recognize them as the same column. As a result, it will not combine the columns. 

So we first make the names of all of the columns lowercase. 

```{r}
names( pc.2021.subset ) <- tolower( names( pc.2021.subset ) )
names( pc.2020.subset ) <- tolower( names( pc.2020.subset ) )
names( pc.2019.subset ) <- tolower( names( pc.2019.subset ) )
names( pf.2021.subset ) <- tolower( names( pf.2021.subset ) )
names( pf.2020.subset ) <- tolower( names( pf.2020.subset ) )
```




Then filter the data to include only the columns we plan to use in the analysis. We use the following fields:


**Form 990-PF Columns**

| Column            | Description                                      | Location in Form 990 |
|-------------------|--------------------------------------------------|----------------------|
| ein               | Employer Identification Number                   | Header               |
| payrolltx         | Payroll taxes                                    | 990 Core_Pt IX-10(A) |
| compnsatncurrofcr | Compensation of current officers, directors, etc | 990 Core_Pt IX-5(A)  |
| compnsatnandothr  | Compensation of disqualified persons             | 990 Core_Pt IX-6(A)  |
| othrsalwages      | Other salaries and wages                         | 990 Core_Pt IX-7(A)  |
| pensionplancontrb | Pension plan contributions                       | 990 Core_Pt IX-8(A)  |
| othremplyeebenef  | Other employee benefits                          | 990 Core_Pt IX-9(A)  |

**Form 990-PF Columns**

| Column         | Description                      | Location in Form 990    |
|----------------|----------------------------------|-------------------------|
| ein            | Employer Identification Number   | Header                  |
| compofficers   | Compensation of officers         | 990-PF Pt I-13, col (a) |
| pensplemplbenf | Pension plans, employee benefits | 990-PF Pt I-15, col (a) |


```{r}
pc.2021.subset2 <- 
  pc.2021.subset %>%
  select( ein, payrolltx, compnsatncurrofcr, 
          compnsatnandothr, othrsalwages, 
          pensionplancontrb, othremplyeebenef ) 

pc.2020.subset2 <- 
  pc.2020.subset %>%
  select( ein, payrolltx, compnsatncurrofcr, 
          compnsatnandothr, othrsalwages, 
          pensionplancontrb, othremplyeebenef ) 

pc.2019.subset2 <- 
  pc.2019.subset %>%
  select( ein, payrolltx, compnsatncurrofcr, 
          compnsatnandothr, othrsalwages, 
          pensionplancontrb, othremplyeebenef ) 

pf.2021.subset2 <- 
  pf.2021.subset %>%
  select( ein, compofficers, pensplemplbenf ) 

pf.2020.subset2 <- 
  pf.2020.subset %>%
  select( ein, compofficers, pensplemplbenf ) 
```


### Stack the data

```{r}
pc <- 
  bind_rows( pc.2021.subset2, 
             pc.2020.subset2, 
             pc.2019.subset2 )
pf <- 
  bind_rows( pf.2021.subset2, 
             pf.2020.subset2 )
```



### Remove duplicate EINs

There might be some duplicate EINs in the combined datasets because some organizations may have submitted amended returns.


```{r, results="hold"}
# roughly 1,200 duplicates
n_distinct( pc$ein )
nrow( pc )

n_distinct( pf$ein )
nrow( pf )
```

Remove duplicate EINs from the combined datasets

```{r, results="hold"}
pc <- pc %>% distinct( ein, .keep_all = TRUE )
pf <- pf %>% distinct( ein, .keep_all = TRUE )

# check to ensure it worked
n_distinct( pc$ein ) == nrow( pc )
n_distinct( pf$ein ) == nrow( pf )
```


## Analysis


### Step 1: Calculate Payroll Taxes Paid by Nonprofits 

Calculate the amount of payroll taxes paid by Form 990 filers in 2019: 

```{r}
# helpful function for printing large numbers
dollarize <- function(x) 
{ paste0( "$", format( round(x,0), big.mark="," ) ) }

prtax <- sum( pc$payrolltx, na.rm = T )
prtax %>% dollarize()
```



### Step 2: Estimate the Effective Payroll Tax Rate 

Calculate a payroll tax/salaries and benefits ratio for Form 990 filers.

To calculate the ratio, we divide the amount of payroll taxes that Form 990 filers pay by the amount that Form 990 filers spend on salaries and benefits.

Calculate the amount that Form 990 filers spend on salaries and benefits.

```{r}
salaries <- 
  sum( pc$compnsatncurrofcr,
       pc$compnsatnandothr,
       pc$othrsalwages,
       pc$pensionplancontrb,
       pc$othremplyeebenef, 
       na.rm = T )
```

The following function can be helpful for printing large numbers: 

```{r}
salaries %>% dollarize()
```


We estimate the effective tax rate by tabulating the amount of payroll taxes paid by nonprofits divided by total salary costs: 

```{r}
ratio <- prtax / salaries
paste0( 100*round( ratio, 4 ), "%" )
```

### Step 3: Estimate PF Contributions 

It's not perfect, but it allows us to roughly estimate the payroll taxes paid by private foundations since that data is not included in the SOI PF Extracts: 

```{r}
salaries.pf <- 
  sum( pf$compofficers, 
       pf$pensplemplbenf, 
       na.rm = T )

salaries.pf %>% dollarize()
```
Estimated payroll taxes contributed by PFs: 

```{r}
prtax.pf <- ratio * salaries.pf

prtax.pf %>% dollarize()
```



### Step 4: Tabulate Totals 

Add the amount of payroll taxes that nonprofits pay to the estimated amount of payroll taxes contributed by Private Foundations: 

```{r}
tax.total <- prtax + prtax.pf
tax.total %>% dollarize()
```



### Step 5: Adjust for Inflation

Normally inflation would be marginal over a 2-3 year period, but inflation was excessive during the time frame of the study (2019 to 2023). As a result we should include an inflation adjustment. This scaling factor can be obtained by entering the start and end year on any reputable online inflation generator. 

```{r}
( tax.total * 1.1885 ) %>% dollarize()
```

We see that the estimate increased from $55 billion to $66 billion, a significant amount. 

We would interpret this number either as the value of 2019 nonprofit payroll tax contributions in today's dollars, or as an estimate of what nonprofits are likely paying in 2023 if we assume employment rates were similar in those two time periods. 



## Conclusion

We estimate that nonprofits pay approximately $66 billion per year in payroll taxes. 

As we described at the beginning of this tutorial, this is a conservative estimate. 

* It does not include the payroll taxes that Form 990-EZ or Form 990-N filers pay. 
* It underestimates the payroll taxes that private foundations pay.
* It also does not include the payroll taxes from nonprofits that are not required to file annual 990 returns (for example, churches). 

This approach is a useful way to demonstrate the size and economic contributions of the nonprofit sector using freely available IRS administrative tax data.





<br>
<br>
<br>
<br>





