---
title: Payroll Taxes V3
date: 2023-08-28
description: This data story outlines the process for estimating the amount of payroll taxes nonprofits pay annually.
featured: true
featuredOrder: 1
type: methods
categories:
  - SOI extracts
  - payroll tax
author:
  - id: hmartin
  - id: jlecy
citation: 
  container-title: National Center for Charitable Statistics
  volume: 1
  issue: 1
  doi: 10.5555/12345678
links:
  - header: Replication Files
    links:  
    - text: Data + Code
      href: https://tinyurl.com/ylydvtj9
      icon: download
    - text: Code
      href: https://raw.githubusercontent.com/UrbanInstitute/nccs-recipes/main/vignettes/payroll.R
      icon: link
  - header: In Context 
    links:
    - text: Nonprofit Infrastructure Coalition 
      href: https://independentsector.org/wp-content/uploads/2023/04/nic-agenda-2023.pdf
      icon: download
---



The nonprofit sector is persistently striving to measure its scale of activities to better understand its impact as well as convey the scale of the work to key stakeholders to highlight societal benefits. Clear ways of measuring activities can help engage policymakers and donors to ensure support for the work is sustained.

How does one measure the size of the nonprofit sector? Researchers and advocates commonly use counts of nonprofits, revenues, assets, and program expenses. The mobilization of human resources, both paid and volunteer, is perhaps a more salient measure of the depth of engagement within communities but is has historically been a harder statistic to generate due to limited reporting requirements (nonprofits report estimates of employees on the 990 forms but they are not required to differentiate part-time and full-time employees and the information has not been systematically digitized until recently). Partnerships with the Quarterly Census of Employment and Wages have helped more precisely measure the size of the nonprofit workforce but estimates are currently only updated every five years. 

A less common but perhaps more accurate approach to measuring employment levels is through payroll tax disclosures. Although most nonprofits are exempt from federal income taxes they are still required to pay federal payroll taxes. Taxes are levied for both part and full-time employees, meaning the aggregate level of payroll taxes can be used to estimate the number of full-time equivalent workers employed by the sector.  

This data story estimates the amount of payroll taxes the nonprofit sector contributes annually, in 2023 dollars.

To replicate our methodology, please see the tutorial at the end.

## Estimating Payroll Tax Contributions

To estimate the payroll taxes that the nonprofit sector contributes annually, we use 2019 IRS 990 data, since it is the most recent full year of available data. IRS Form 990 data extracts, available at <https://www.irs.gov/statistics/soi-tax-stats-annual-extract-of-tax-exempt-organization-financial-data>. 

We use 3 years of data extracts since there can be a 2-year lag in returns being filed, so we first compile all of the 2019 990 returns from across the 2019, 2020, and 2021 SOI extracts. 

### Form 990 Data

The Internal Revenue Service (IRS) requires tax-exempt nonprofits to file annual returns.

* Private foundations must file Form 990-PF.
* Most nonprofits must file Form 990, 990-EZ, or 990-N.
    + Those with gross receipts >= $200K or total assets >= $500K must file Form 990.
    + Those with gross receipts <$200K and total assets <$500K can file Form 990-EZ.
    + Those with gross receipts <=$50K can file Form 990-N.
* Some nonprofits, such as churches, do not have to file annual returns. See <https://www.irs.gov/charities-non-profits/annual-exempt-organization-return-who-must-file>.

### Data Limitations 

What does Form 990 series tell us about the amount of payroll taxes that nonprofits pay?

***Larger Nonprofits:*** The Form 990 collects data on the amount of payroll taxes that larger nonprofits (i.e., those with gross receipts >= $200K or total assets >= $500K) contribute (**$55B in 2019**).

***Private foundations:*** The Form 990-PF does not collect data on the amount of payroll taxes that private foundations contribute However, it does collect data on the amount that private foundations spend on salaries and benefits ($3B). The Form 990 collects this data for larger nonprofits, as well. Therefore, we can divide the amount of payroll taxes that larger nonprofits contribute by the amount that larger nonprofits spend on payroll taxes to calculate a payroll tax/salaries and benefits ratio (5.75%). We can then apply that ratio to the salaries and benefits data from the Form 990-PF to estimate the amount of payroll taxes that private foundations contribute.

However, the IRS's Form 990-PF data extract only includes the columns with data on "compensation of officers, directors, trustees, etc." and "pension plans, employee benefits;" it excludes the column with data on "other employee salaries and wages." Therefore, we will underestimate the amount of payroll taxes that private foundations contribute. Additionally, the IRS does not provide a data extract for the 2019 Form 990-PF file. This creates a further underestimation of the amount of payroll taxes that private foundations contribute (**estimated $153M in 2019**).

***Smaller Nonprofits:*** The Form 990-EZ does not collect data on the amount of payroll taxes that smaller nonprofits (i.e., those with gross receipts <$200K and total assets <$500K) contribute It does collect data on the amount that smaller nonprofits spend on salaries and benefits, but the IRS's Form 990-EZ data extracts exclude these data. Therefore, we are unable to estimate the amount of payroll taxes that smaller nonprofits contribute.

***Invisible Nonprofits:*** The Form 990-N does not collect data on the amount of payroll taxes that the smallest nonprofits (i.e., those with gross receipts <=$50K) contribute, nor does it collect data on the amount that these nonprofits spend on salaries and benefits. Therefore, we are unable to estimate the amount of payroll taxes that the smallest nonprofits contribute.

We are also unable to estimate the amount of payroll taxes that nonprofits that are not required to file annual returns contribute.

***Total:*** Adding the amount of payroll taxes that larger nonprofits contributed in 2019 to the estimated amount of payroll taxes that private foundations contributed in 2019 and adjusting 18.85% for inflation (per <https://salaryinflation.com/>), **we estimate that the nonprofit sector contributes $66B annually in payroll taxes.** Although this is a conservative estimate of payroll taxes, it demonstrates the huge size and economic contributions of the nonprofit sector.

## Data Vignette 




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
```



## Setup 


### Packages 

You can install R packages as follows:

```{r, eval=F}
install.packages( "knitr" )
install.packages( "tidyverse" )
install.packages( "pander" )
```

If you already have these packages, you only need to load the libraries.

```{r}
library( tidyverse )  # data wrangling
library( knitr )      # pretty formats
library( pander )     # pretty formats
```


### Data 

Load the SOI Extract data obtained from the IRS: 

```{r, echo=F, eval=T, cache=T}
url.pc.2019 <- "C:/Users/jdlec/Downloads/payroll/19eoextract990.csv"
url.pc.2020 <- "C:/Users/jdlec/Downloads/payroll/20eoextract990.csv"
url.pc.2021 <- "C:/Users/jdlec/Downloads/payroll/21eoextract990.csv"
url.pf.2020 <- "C:/Users/jdlec/Downloads/payroll/20eoextract990pf.csv"
url.pf.2021 <- "C:/Users/jdlec/Downloads/payroll/21eoextract990pf.csv"

pc.2019 <- readr::read_csv( url.pc.2019 )
pc.2020 <- readr::read_csv( url.pc.2020 )
pc.2021 <- readr::read_csv( url.pc.2021 )
pf.2020 <- readr::read_csv( url.pf.2020 )
pf.2021 <- readr::read_csv( url.pf.2021 )
```

Download all data and read locally: 

```{r, eval=F}
URL <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/payroll.zip"
download.file( URL, destfile="payroll.zip" )
unzip( "payroll.zip" )

pc.2019 <- readr::read_csv( "payroll/19eoextract990.csv" )
pc.2020 <- readr::read_csv( "payroll/20eoextract990.csv" )
pc.2021 <- readr::read_csv( "payroll/21eoextract990.csv" )
pf.2020 <- readr::read_csv( "payroll/20eoextract990pf.csv" )
pf.2021 <- readr::read_csv( "payroll/21eoextract990pf.csv" )
```

Or read directly from the URLs: 

```{r, eval=F}
url.pc.2019 <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/19eoextract990.csv"
url.pc.2020 <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/20eoextract990.csv"
url.pc.2021 <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/21eoextract990.csv"
url.pf.2020 <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/20eoextract990pf.csv"
url.pf.2021 <- "https://nccsdata.s3.us-east-1.amazonaws.com/replication/payroll/21eoextract990pf.csv"

pc.2019 <- readr::read_csv( url.pc.2019 )
pc.2020 <- readr::read_csv( url.pc.2020 )
pc.2021 <- readr::read_csv( url.pc.2021 )
pf.2020 <- readr::read_csv( url.pf.2020 )
pf.2021 <- readr::read_csv( url.pf.2021 )
```



Preview the data: 

```{r}
head( pc.2019[,1:6] ) %>% pander()
```


The data dictionaries are available in the zipped directory. Additional information can be obtained at the [SOI Extracts page](https://www.irs.gov/statistics/soi-tax-stats-annual-extract-of-tax-exempt-organization-financial-data).


## Data Preparation 

### Variable Cleanup

Convert six digit tax period dates (YYYYMM) to four digit years (YYYY): 




```{r}
pc.2021$year <- pc.2021$tax_pd  %>% substr( 1, 4 )
pc.2020$year <- pc.2020$tax_pd  %>% substr( 1, 4 )
pc.2019$year <- pc.2019$tax_pd  %>% substr( 1, 4 )
pf.2020$year <- pf.2020$TAX_PRD %>% substr( 1, 4 )
pf.2021$year <- pf.2021$TAX_PRD %>% substr( 1, 4 )
```

```{r, results="hold"}
head(pc.2021$tax_pd)
head(pc.2021$year)
```

### Isolate the Study Period

SOI Extract files are organized by filing dates, the dates the returns were received by the IRS. The filing date is usually around six months after the end of an organization's fiscal year, which is also twelve months ahead of the calendar year that correspond to the nonprofits activities. 

In addition, nonprofits can back-file late returns or submit amended returns, so although the year of the SOI extract file is different from the periods of data contained within: 

```{r}
pc.2021 %>%
  group_by( year ) %>%
  summarise( n() ) %>%
  kable( col.names = c( "Tax Period", "Frequency" ),
         caption = "Tax Periods in the 2021 Form 990 Dataset" )
```


```{r}
pf.2021 %>%
  group_by( year ) %>%
  summarise( n() ) %>%
  kable( col.names = c( "Tax Period", "Frequency" ),
         caption = "Tax Periods in the 2021 Form 990-PF Dataset" )
```


We need data that corresponds to the 2019 calendar year, the most recent full year available at the time of writing. 

```{r}
pc.2021.subset <- filter( pc.2021, year=="2019" )
pf.2021.subset <- filter( pf.2021, year=="2019" )
pc.2020.subset <- filter( pc.2020, year=="2019" )
pf.2020.subset <- filter( pf.2020, year=="2019" )
pc.2019.subset <- filter( pc.2019, year=="2019" )
```

Inspect to ensure we have the correct data now: 

```{r}
pc.2021.subset %>%
  group_by( year ) %>%
  summarise( n() ) %>%
  kable( col.names = c( "Tax Period", "Frequency" ),
         caption = "Tax Periods in the Subsetted 2021 Form 990 Dataset" )

pf.2021.subset %>%
  group_by( year ) %>%
  summarise( n() ) %>%
  kable( col.names = c( "Tax Period", "Frequency" ),
         caption = "Tax Periods in the Subsetted 2021 Form 990-PF Dataset" )
```


### Harmonizing the Data

In order to combine files we need to ensure that the variable names are the same. 

R is case-sensitive; if one dataset uses uppercase letters for a column name and another dataset uses lowercase letters for a column name, it will not recognize them as the same column. As a result, it will not combine the columns. 

So we first make the names of all of the columns lowercase. 

```{r}
names( pc.2021.subset ) <- tolower( names( pc.2021.subset ) )
names( pc.2020.subset ) <- tolower( names( pc.2020.subset ) )
names( pc.2019.subset ) <- tolower( names( pc.2019.subset ) )
names( pf.2021.subset ) <- tolower( names( pf.2021.subset ) )
names( pf.2020.subset ) <- tolower( names( pf.2020.subset ) )
```




Then filter the data to include only the columns we plan to use in the analysis. We use the following fields:


**Form 990-PF Columns**

| Column            | Description                                      | Location in Form 990 |
|-------------------|--------------------------------------------------|----------------------|
| ein               | Employer Identification Number                   | Header               |
| payrolltx         | Payroll taxes                                    | 990 Core_Pt IX-10(A) |
| compnsatncurrofcr | Compensation of current officers, directors, etc | 990 Core_Pt IX-5(A)  |
| compnsatnandothr  | Compensation of disqualified persons             | 990 Core_Pt IX-6(A)  |
| othrsalwages      | Other salaries and wages                         | 990 Core_Pt IX-7(A)  |
| pensionplancontrb | Pension plan contributions                       | 990 Core_Pt IX-8(A)  |
| othremplyeebenef  | Other employee benefits                          | 990 Core_Pt IX-9(A)  |

**Form 990-PF Columns**

| Column         | Description                      | Location in Form 990    |
|----------------|----------------------------------|-------------------------|
| ein            | Employer Identification Number   | Header                  |
| compofficers   | Compensation of officers         | 990-PF Pt I-13, col (a) |
| pensplemplbenf | Pension plans, employee benefits | 990-PF Pt I-15, col (a) |


```{r}
pc.2021.subset2 <- 
  pc.2021.subset %>%
  select( ein, payrolltx, compnsatncurrofcr, 
          compnsatnandothr, othrsalwages, 
          pensionplancontrb, othremplyeebenef ) 

pc.2020.subset2 <- 
  pc.2020.subset %>%
  select( ein, payrolltx, compnsatncurrofcr, 
          compnsatnandothr, othrsalwages, 
          pensionplancontrb, othremplyeebenef ) 

pc.2019.subset2 <- 
  pc.2019.subset %>%
  select( ein, payrolltx, compnsatncurrofcr, 
          compnsatnandothr, othrsalwages, 
          pensionplancontrb, othremplyeebenef ) 

pf.2021.subset2 <- 
  pf.2021.subset %>%
  select( ein, compofficers, pensplemplbenf ) 

pf.2020.subset2 <- 
  pf.2020.subset %>%
  select( ein, compofficers, pensplemplbenf ) 
```


### Stack the data

```{r}
pc <- 
  bind_rows( pc.2021.subset2, 
             pc.2020.subset2, 
             pc.2019.subset2 )
pf <- 
  bind_rows( pf.2021.subset2, 
             pf.2020.subset2 )
```



### Remove duplicate EINs

There might be some duplicate EINs in the combined datasets because some organizations may have submitted amended returns.


```{r}
# roughly 1,200 duplicates
n_distinct( pc$ein )
nrow( pc )

n_distinct( pf$ein )
nrow( pf )
```

Remove duplicate EINs from the combined datasets

```{r, results="hold"}
pc <- pc %>% distinct( ein, .keep_all = TRUE )
pf <- pf %>% distinct( ein, .keep_all = TRUE )

# check to ensure it worked
n_distinct( pc$ein ) == nrow( pc )
n_distinct( pf$ein ) == nrow( pf )
```


## Analysis


### Step 1: Calculate Payroll Taxes Paid by Nonprofits 

Calculate the amount of payroll taxes paid by Form 990 filers in 2019: 

```{r}
# helpful function for printing large numbers
dollarize <- function(x) 
{ paste0( "$", format( round(x,0), big.mark="," ) ) }

prtax <- sum( pc$payrolltx, na.rm = T )
prtax %>% dollarize()
```



### Step 2: Estimate the Effective Payroll Tax Rate 

Calculate a payroll tax/salaries and benefits ratio for Form 990 filers.

To calculate the ratio, we divide the amount of payroll taxes that Form 990 filers pay by the amount that Form 990 filers spend on salaries and benefits.

Calculate the amount that Form 990 filers spend on salaries and benefits.

```{r}
salaries <- 
  sum( pc$compnsatncurrofcr,
       pc$compnsatnandothr,
       pc$othrsalwages,
       pc$pensionplancontrb,
       pc$othremplyeebenef, 
       na.rm = T )
```

The following function can be helpful for printing large numbers: 

```{r}
salaries %>% dollarize()
```


We estimate the effective tax rate by tabulating the amount of payroll taxes paid by nonprofits divided by total salary costs: 

```{r}
ratio <- prtax / salaries
paste0( 100*round( ratio, 4 ), "%" )
```

### Step 3: Estimate PF Contributions 

It's not perfect, but it allows us to roughly estimate the payroll taxes paid by private foundations since that data is not included in the SOI PF Extracts: 

```{r}
salaries.pf <- 
  sum( pf$compofficers, 
       pf$pensplemplbenf, 
       na.rm = T )

salaries.pf %>% dollarize()
```
Estimated payroll taxes contributed by PFs: 

```{r}
prtax.pf <- ratio * salaries.pf

prtax.pf %>% dollarize()
```



### Step 4: Tabulate Totals 

Add the amount of payroll taxes that nonprofits pay to the estimated amount of payroll taxes contributed by Private Foundations: 

```{r}
tax.total <- prtax + prtax.pf
tax.total %>% dollarize()
```



### Step 5: Adjust for Inflation

Normally inflation would be marginal over a 2-3 year period, but inflation was excessive during the time frame of the study (2019 to 2023). As a result we should include an inflation adjustment. This scaling factor can be obtained by entering the start and end year on any reputable online inflation generator. 

```{r}
( tax.total * 1.1885 ) %>% dollarize()
```

We see that the estimate increased from $55 billion to $66 billion, a significant amount. 

We would interpret this number either as the value of 2019 nonprofit payroll tax contributions in today's dollars, or as an estimate of what nonprofits are likely paying in 2023 if we assume employment rates were similar in those two time periods. 



## Conclusion

We estimate that nonprofits pay approximately $66 billion per year in payroll taxes. 

As we described at the beginning of this tutorial, this is a conservative estimate. 

* It does not include the payroll taxes that Form 990-EZ or Form 990-N filers pay. 
* It underestimates the payroll taxes that private foundations pay.
* It also does not include the payroll taxes from nonprofits that are not required to file annual 990 returns (for example, churches). 

This approach is a useful way to demonstrate the size and economic contributions of the nonprofit sector using freely available IRS administrative tax data.


<br>
<br>
<hr>
<br>
<br>







<br>
<br>
<hr>
<br>
<br>





