---
title: "NCCS Revocations Data Catalog"
format: html
editor: visual
execute: 
  echo: false
  warning: false
---

```{r}
#| echo: false

library( dplyr )
library( knitr )
library( kableExtra )
library( stringr )
library( flextable )
library( pander )

source("build-catalog-functions.R")

fn  <- "AWS-NCCSDATA.csv"
d   <- read.csv( fn )

# fn   <- "AWS-NCCSDATA.csv"
# path <- "https://raw.githubusercontent.com/UrbanInstitute/nccs/main/catalogs/"
# d    <- read.csv( paste0( path, fn  ) )

```

<br>
<hr>
<br>


#### Import to R 

Retrieve the most recent file available using the following script: 

```{r, eval=F, echo=T}
part_1 <- "https://nccsdata.s3.us-east-1.amazonaws.com/raw/revocations/" 
YYYYMM <- format( Sys.Date(), "%Y-%m" )
part_2 <- "-REVOCATIONS-ORGS.csv"

URL <- paste0( part_1, YYYYMM, part_2 )
d <- read.csv( URL )
```


<br>
<hr>
<br>

#### Download Data in CSV Format 

```{r}
catalog <- get_catalog_revocations( d$URL )

catalog %>%
  knitr::kable( format="html", escape=FALSE ) %>%
  kableExtra::kable_styling() 
```

<br>

<hr>

<br>

<style>


h1.title {
 font-size: 60px;
 color: #0a4c6a;
}


h4, .h4 {
    font-size: 40px;
    color: #1696d2;
    margin-bottom: 50px;
}



.button {
  background-color: white;
  color:  #12719e;
  border: 2px solid  #12719e;   /* #008CBA; */
  padding: 5px 10px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 14px;
  border-radius: 12px;
  width: 150px;
}
 
.button {
  transition-duration: 0.4s;
}

.button:hover {
  background-color: #1696d2; 
  color: white;
  border: 2px solid #12719e;
}


.button2 {
  background-color: white;
  color: #12719e;
  border: 2px solid #9d9d9d;   /* #008CBA; */
  padding: 5px 10px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 14px;
  border-radius: 12px;
  width: 150px;
}

.button2 {
  transition-duration: 0.4s;
}

.button2:hover {
  background-color: #fff2cf; 
  color: #1696d2;
  font-weight: bold;
  border: 2px solid #1696d2;
}


.center {text-align: center;}

table td {
  vertical-align: middle;
}

blockquote {
    border-left: 4px solid #1696d2;
    margin-left: 10px;
    margin-inline-start: 0px;
    margin-inline-end: 0px;
    padding: 15px;
}

.button3 {
  background-color: #0a4c6a;
  color: white;
  border: 2px solid #0a4c6a;   
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 16px;
  border-radius: 4px;
  width: 100px;
}
</style>
