---
title: "Core Data Catalog"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| echo: false

suppressPackageStartupMessages({
  library(dplyr)
  library(knitr)
  library(kableExtra)
  library(stringr)
})

source("R/core_helpers.R")

manifest <- read.csv("AWS-NCCSDATA.csv", stringsAsFactors = FALSE)

render_year_table <- function(df) {
  knitr::kable(
    df,
    escape    = FALSE,
    col.names = c("Year", "Download", "Dictionary", "Quality Report", "Size"),
    format    = "html",
    row.names = FALSE
  ) |>
    kableExtra::kable_styling()
}
```

The NCCS Core Data Series is a panel of nonprofit organizations derived from annual IRS Form 990 filings. This catalog organizes the published files by **form type** — pick the variant whose filing scope matches your analysis.

For background on what the Core Series contains and its scope limitations, see the [Core overview page](../datasets/core.html).

::: {.callout-note title="Files are organized by tax year, not filing year"}
Every year label in this catalog refers to the **tax year covered by the return** — the fiscal period the filing describes — not the calendar year the IRS received it. A nonprofit's 2022 tax-year return is in the `2022` directory regardless of when it was actually filed (typically 6&ndash;18 months later). Recent tax years are still in-progress as the IRS continues to release returns over time.
:::

## Which dataset should I use?

| If you need... | Use | Why |
|:---|:---|:---|
| Full Form 990 filers only, max variables | **990** | Larger nonprofits filing the long form (251 columns in 2024). Best for detailed financial analysis. |
| Short-form filers only | **990EZ** | Smaller nonprofits filing the short form (76 columns in 2024). |
| Every 990 + 990EZ filer in one file | **990 Combined** | The broadest population per year. Stacks 990 and 990EZ filers using the 53-column shared variable subset (plus pipeline-derived universals: `tax_year`, `tax_month`, `is_501c3`, `extract_year`, `is_amendment`). |
| Private foundations | **990PF** | Filed by all private foundations regardless of size. Separate variable set from public charities (187 columns in 2024). |

Each file pools 501(c)(3) charities and all other 501(c) types into a single product. Use the **`is_501c3`** column (boolean, `TRUE` for 501(c)(3) public charities) to filter when needed; the raw **`subsection_cd`** (IRC EO subsection code, 1&ndash;93) is also available for finer 501(c) subclassification.

<br><hr><br>

## 990 {#section-990}

Full Form 990 filers — the long-form return required for nonprofits with gross receipts ≥ \$200k or assets ≥ \$500k. The richest variable set in the Core Series (251 columns in 2024) but a narrower population than 990 Combined.

### Recent Releases

```{r}
recent <- build_core_year_section(manifest, "990", n_recent = 5)
render_year_table(recent)
```

### Browse Older Releases

<details>
<summary><strong>Show all 990 releases</strong></summary>

```{r}
all_990 <- build_core_year_section(manifest, "990", n_recent = NA)
render_year_table(all_990)
```

</details>

### Access via S3

```
s3://nccsdata/processed/core/{YYYY}/990/
```

<br><hr><br>

## 990EZ {#section-990ez}

Short-form filers only — Form 990EZ is filed by mid-sized nonprofits (gross receipts < \$200k and assets < \$500k). Fewer columns than the full 990 (76 in 2024) but covers a substantially different population.

Most users want **990 Combined** below; this standalone slice is useful when you specifically need only the EZ filers.

### Recent Releases

```{r}
recent_ez <- build_core_year_section(manifest, "990ez", n_recent = 5)
render_year_table(recent_ez)
```

### Browse Older Releases

<details>
<summary><strong>Show all 990EZ releases</strong></summary>

```{r}
all_ez <- build_core_year_section(manifest, "990ez", n_recent = NA)
render_year_table(all_ez)
```

</details>

### Access via S3

```
s3://nccsdata/processed/core/{YYYY}/990ez/
```

<br><hr><br>

## 990 Combined {#section-990combined}

The broadest single-file product per year — Form 990 + Form 990EZ filers stacked together using the 53-column shared variable subset (plus pipeline-derived universals: `tax_year`, `tax_month`, `is_501c3`, `extract_year`, `is_amendment`). Use this when you want the full population of 990/990EZ filers and don't need the long-form-only variables.

### Recent Releases

```{r}
recent_c <- build_core_year_section(manifest, "990combined", n_recent = 5)
render_year_table(recent_c)
```

### Browse Older Releases

<details>
<summary><strong>Show all 990 Combined releases</strong></summary>

```{r}
all_c <- build_core_year_section(manifest, "990combined", n_recent = NA)
render_year_table(all_c)
```

</details>

### Access via S3

```
s3://nccsdata/processed/core/{YYYY}/990combined/
```

<br><hr><br>

## 990PF {#section-990pf}

Private foundations — all private foundations must file Form 990PF regardless of size. Has its own variable set distinct from public charities (187 columns in 2024).

### Recent Releases

```{r}
recent_pf <- build_core_year_section(manifest, "990pf", n_recent = 5)
render_year_table(recent_pf)
```

### Browse Older Releases

<details>
<summary><strong>Show all 990PF releases</strong></summary>

```{r}
all_pf <- build_core_year_section(manifest, "990pf", n_recent = NA)
render_year_table(all_pf)
```

</details>

### Access via S3

```
s3://nccsdata/processed/core/{YYYY}/990pf/
```

<br><hr><br>

## File Conventions

Every year/form directory contains two artifacts on S3:

```
processed/core/{YYYY}/{FORM_TYPE}/
    core_{YYYY}_{FORM_TYPE}.csv             # the data
    core_{YYYY}_{FORM_TYPE}_dictionary.csv  # variable definitions for this year
```

Quality reports are published on GitHub Pages alongside the [CORE Pipeline Guide](https://urbaninstitute.github.io/nccs-data-core/) and linked from the Quality Report column in each table above. The [Quality Reports Index](https://urbaninstitute.github.io/nccs-data-core/12-quality-reports.html) lists all 109 published reports.

Data dictionaries are per-year because the variable set evolves as the underlying 990 forms change. Always read the dictionary for the year you're using rather than assuming column stability across years.

Filter on **`is_501c3`** (boolean) to subset to 501(c)(3) public charities or their complement. The raw **`subsection_cd`** (IRC EO subsection code, 1&ndash;93) is also present for finer breakdowns.

Tables above start in 2010 for the 990/990EZ/990 Combined forms and in 2010 for 990PF — earlier years exist as placeholder files only. For pre-2010 coverage, use the **Legacy: Harmonized Core** section below.

<br><hr><br>

## Before you analyze

Known upstream-IRS data shapes that look like pipeline bugs but aren't &mdash; documented in the pipeline guide so analysts don't have to rediscover them:

- [No top-coding or winsorization](https://urbaninstitute.github.io/nccs-data-core/08-output-schema.html#no-top-coding-or-winsorization) &mdash; identical maxes across related columns (e.g. `total_contributions` = `total_revenue`) reflect real-world equality for one filer, not capping.
- [~12% of 990-EZ rows fail the balance-sheet identity](https://urbaninstitute.github.io/nccs-data-core/11-upstream-source-quirks.html#filer-supplied-balance-sheet-identity-failures-12-of-990-ez-rows) &mdash; filer-supplied, not a harmonization error.
- [Schedule A `_sec170` / `_sec509` zeros are filer-test-specific](https://urbaninstitute.github.io/nccs-data-core/11-upstream-source-quirks.html#schedule-a-sec170-sec509-blocks-are-filer-test-specific-zeros-are-not-necessarily-missing) &mdash; cross-reference `non_pf_status_reason` before interpreting them as missing data.
- [Filer-supplied negatives on revenue fields](https://urbaninstitute.github.io/nccs-data-core/11-upstream-source-quirks.html#filer-supplied-negatives-on-conceptually-non-negative-fields) &mdash; IRS-supplied as-filed; not amendments.

Full pipeline guide and per-form quality reports are at the [CORE Pipeline Guide](https://urbaninstitute.github.io/nccs-data-core/).

<br><hr><br>

## Blank IRS Forms Archive

Blank IRS Form 990, 990-EZ, 990-PF, their schedules (A&ndash;O and R), and the corresponding instruction booklets are mirrored at `s3://nccsdata/raw/core/forms/`. Each cell below links the blank **Form** and (where IRS publishes one standalone) the **Instr.**uction booklet. Empty cells reflect years IRS did not publish that combination &mdash; not a coverage gap on our side. See the [Forms archive](https://urbaninstitute.github.io/nccs-data-core/02-data-lineage.html) section of the Pipeline Guide for the year-coverage details and refresh SOP.

Direct download URLs follow:

```
https://nccsdata.s3.us-east-1.amazonaws.com/raw/core/forms/<basename>_<YYYY>.pdf
```

A note on schedule instructions: M, N, and parts of I/L have no standalone instruction PDF because IRS folds those instructions into the main `i990_YYYY.pdf` booklet linked under Form 990 below.

### Main Forms

```{r}
forms_main <- build_forms_matrix(manifest, CORE_MAIN_FORMS)
knitr::kable(
  forms_main,
  escape    = FALSE,
  format    = "html",
  row.names = FALSE,
  align     = c("l", rep("c", length(CORE_MAIN_FORMS)))
) |>
  kableExtra::kable_styling(full_width = FALSE)
```

### Schedules

```{r}
#| results: asis
cat("<div style='overflow-x: auto;'>")
forms_sched <- build_forms_matrix(manifest, CORE_SCHEDULES)
cat(
  knitr::kable(
    forms_sched,
    escape    = FALSE,
    format    = "html",
    row.names = FALSE,
    align     = c("l", rep("c", length(CORE_SCHEDULES)))
  ) |>
    kableExtra::kable_styling(full_width = FALSE)
)
cat("</div>")
```

<br><hr><br>

## Legacy: Harmonized Core (deprecated)

<details>
<summary><strong>Show legacy <code>harmonized/core/</code> files</strong> (for reproducibility only — use the by-form-type tables above)</summary>

These files predate the current by-form-type layout. They split each year's filers along two axes — **tax-exempt class** (501c3 charities vs. all other 501c types) and **form scope** (PC = full 990 only, PZ = 990 + 990EZ) — producing five products. The current pipeline pools these into single per-form-type files with an `is_501c3` column for filtering, so the products below are kept only to support replication of analyses built on the older layout.

**Shared dictionaries** (PC/PZ variants share a single dictionary; PF has its own):

- [PC/PZ data dictionary](https://nccsdata.s3.amazonaws.com/harmonized/core/CORE-HRMN_dd.csv)
- [PF data dictionary](https://nccsdata.s3.amazonaws.com/harmonized/core/501c3-pf/dd/DD-PF-HRMN-V0.csv)

### 501(c)(3) Charities &mdash; PC (full 990 only)

```{r}
render_legacy_table <- function(df) {
  knitr::kable(
    df,
    escape    = FALSE,
    col.names = c("Year", "Download", "Size"),
    format    = "html",
    row.names = FALSE
  ) |>
    kableExtra::kable_styling()
}
render_legacy_table(build_legacy_core_section(manifest, "501c3-pc"))
```

### 501(c)(3) Charities &mdash; PZ (990 + 990EZ)

```{r}
render_legacy_table(build_legacy_core_section(manifest, "501c3-pz"))
```

### Other 501(c) Nonprofits &mdash; PC (full 990 only)

```{r}
render_legacy_table(build_legacy_core_section(manifest, "501ce-pc"))
```

### Other 501(c) Nonprofits &mdash; PZ (990 + 990EZ)

```{r}
render_legacy_table(build_legacy_core_section(manifest, "501ce-pz"))
```

### Private Foundations &mdash; PF

```{r}
render_legacy_table(build_legacy_core_section(manifest, "501c3-pf"))
```

### Access via S3

```
s3://nccsdata/harmonized/core/{PRODUCT}/marts/
```

where `{PRODUCT}` is one of `501c3-pc`, `501c3-pz`, `501ce-pc`, `501ce-pz`, `501c3-pf`.

</details>

<br><hr><br>

## Documentation & Support

- [Core overview page](../datasets/core.html) — what's in the Core Series, scope limitations, and merging guidance
- [Quality Reports Index](https://urbaninstitute.github.io/nccs-data-core/12-quality-reports.html) — every per-year, per-form quality report
- [BMF Data Catalog](catalog-bmf.html) — Business Master File, for merging organizational attributes (NTEE, 501c type, address) onto Core rows via EIN
- [Census Crosswalk Catalog](catalog-census_crosswalk.html) — aggregate Core data to tracts, counties, or MSAs
- [Contact](https://nccs.urban.org/nccs/contact/) — questions, corrections, data requests

```{r}
#| results: asis
cat(sprintf("\n<p style='font-size: 0.85em; color: #666; margin-top: 2em;'>Last verified: %s</p>\n",
            format(Sys.Date(), "%Y-%m-%d")))
```
