---
title: "BMF Data Catalog"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| echo: false

suppressPackageStartupMessages({
  library(dplyr)
  library(knitr)
  library(kableExtra)
  library(stringr)
})

source("R/bmf_helpers.R")

manifest <- read.csv("AWS-BMF.csv", stringsAsFactors = FALSE)

state_mapping <- c(
  # 50 states + DC
  "AL" = "Alabama",          "AK" = "Alaska",
  "AZ" = "Arizona",          "AR" = "Arkansas",
  "CA" = "California",       "CO" = "Colorado",
  "CT" = "Connecticut",      "DE" = "Delaware",
  "DC" = "District of Columbia",
  "FL" = "Florida",          "GA" = "Georgia",
  "HI" = "Hawaii",           "ID" = "Idaho",
  "IL" = "Illinois",         "IN" = "Indiana",
  "IA" = "Iowa",             "KS" = "Kansas",
  "KY" = "Kentucky",         "LA" = "Louisiana",
  "ME" = "Maine",            "MD" = "Maryland",
  "MA" = "Massachusetts",    "MI" = "Michigan",
  "MN" = "Minnesota",        "MS" = "Mississippi",
  "MO" = "Missouri",         "MT" = "Montana",
  "NE" = "Nebraska",         "NV" = "Nevada",
  "NH" = "New Hampshire",    "NJ" = "New Jersey",
  "NM" = "New Mexico",       "NY" = "New York",
  "NC" = "North Carolina",   "ND" = "North Dakota",
  "OH" = "Ohio",             "OK" = "Oklahoma",
  "OR" = "Oregon",           "PA" = "Pennsylvania",
  "RI" = "Rhode Island",     "SC" = "South Carolina",
  "SD" = "South Dakota",     "TN" = "Tennessee",
  "TX" = "Texas",            "UT" = "Utah",
  "VT" = "Vermont",          "VA" = "Virginia",
  "WA" = "Washington",       "WV" = "West Virginia",
  "WI" = "Wisconsin",        "WY" = "Wyoming",
  # U.S. territories
  "AS" = "American Samoa",   "GU" = "Guam",
  "MP" = "Northern Mariana Islands",
  "PR" = "Puerto Rico",      "VI" = "U.S. Virgin Islands",
  # APO/FPO military
  "AA" = "Armed Forces Americas",
  "AE" = "Armed Forces Europe",
  "AP" = "Armed Forces Pacific",
  # Compact of Free Association
  "FM" = "Federated States of Micronesia",
  "MH" = "Marshall Islands",
  "PW" = "Palau",
  # Missing/unknown state
  "ZZ" = "Unknown / unmapped"
)
```

The IRS Business Master File (BMF) is the authoritative registry of tax-exempt organizations in the United States. NCCS publishes BMF data in several forms; this catalog organizes them by what you're trying to do.

For background on what the BMF contains and its scope limitations, see the [BMF overview page](../datasets/bmf.html).

## Which dataset should I use?

| If you need... | Use | Why |
|:---|:---|:---|
| One row per nonprofit, ever (1989–present) | **Master BMF** | Combines every vintage into a single longitudinal frame |
| A specific month's snapshot (2023-06 or later) | **Transformed monthly BMF** | Cleaned and standardized; column schema is stable across months |
| A specific month's snapshot (1989–2022) | **Harmonized legacy BMF** | Same column schema as transformed; aligns historical vintages |
| The unmodified IRS file | **Raw monthly archive** | Untouched IRS releases, for replication of upstream analysis |
| The original NCCS legacy file | **Raw legacy archive** | Original 501CX-NONPROFIT-PX vintages, for reproducibility only |
| Canonical county FIPS, or to roll up to metro areas | **Geography crosswalks** | Optional join layers for the geocoded BMF (TIGER 2023 / OMB 2023) |

<br><hr><br>

## Master BMF

One row per EIN spanning 1989 to present, with most-recent values for each organization plus `first_vintage_ym`, `last_vintage_ym`, and `bmf_source` to track provenance. This is the primary product for longitudinal analysis and as a sampling frame.

The **geocoded** variant is the recommended default — it carries the same row schema as the plain file plus lat/lon joined against a deduplicated address table.

```{r}
headline <- build_master_headline_table(manifest)
if (!is.null(headline)) {
  knitr::kable(
    headline,
    escape    = FALSE,
    col.names = c("Variant", "Download", "Dictionary", "Quality Report", "Size"),
    format    = "html",
    row.names = FALSE
  ) |>
    kableExtra::kable_styling()
} else {
  cat("_Master BMF is not yet listed in the manifest._\n")
}
```

### Download by State/Territory

For users who only need a single state, the geocoded Master BMF is also published as per-state marts (50 states + DC, U.S. territories, APO/FPO military codes, Compact of Free Association nations, and a `ZZ` bucket for rows with no resolvable state). Each mart carries lat/lon and shares a schema and dictionary with the geocoded headline file above.

```{r}
master <- build_master_section(manifest, state_mapping)
knitr::kable(
  master,
  escape    = FALSE,
  col.names = c("Download Link", "Size", "State/Territory"),
  format    = "html",
  row.names = FALSE
) |>
  kableExtra::kable_styling()
```

### Access via S3

```
s3://nccsdata/geocoding/bmf-master/merged/   # Geocoded variant (recommended)
s3://nccsdata/master/bmf/                    # Plain Master BMF
s3://nccsdata/master/bmf/state_marts/csv/    # Per-state geocoded marts
```

<br><hr><br>

## Monthly BMF

Cleaned, schema-stable monthly snapshots covering 1989 to the present. The transformed pipeline (2023-06+) and the harmonized legacy pipeline (1989-2022) write to different S3 prefixes but share a column schema, so you can stack them across the era boundary without reshaping.

### Recent Releases

One row per release. The Download column links the CSV (canonical format); a parquet variant is also available for some months at the same S3 prefix.

```{r}
recent <- build_combined_monthly_section(manifest, n_recent = 5)
knitr::kable(
  recent,
  escape    = FALSE,
  col.names = c("Year", "Month", "Pipeline", "Download", "Dictionary", "Quality Report", "Size"),
  format    = "html",
  row.names = FALSE
) |>
  kableExtra::kable_styling()
```

### Browse Older Releases

The complete archive of every monthly release — data, dictionary, and quality report — is below. (There is no separate centralized index; this catalog is the canonical listing.)

<details>
<summary><strong>Show all monthly releases (1989–present)</strong></summary>

```{r}
all_releases <- build_combined_monthly_section(manifest, n_recent = NA)
knitr::kable(
  all_releases,
  escape    = FALSE,
  col.names = c("Year", "Month", "Pipeline", "Download", "Dictionary", "Quality Report", "Size"),
  format    = "html",
  row.names = FALSE
) |>
  kableExtra::kable_styling()
```

</details>

### Access via S3

```
s3://nccsdata/processed/bmf/{YYYY_MM}/         # Transformed (2023-06 -> present)
s3://nccsdata/processed/bmf-legacy/{YYYY_MM}/  # Harmonized legacy (1989-06 -> 2022)
```

Each month directory contains:

- `bmf_{YYYY_MM}_processed.csv` (or `.parquet`) — the data
- `bmf_{YYYY_MM}_data_dictionary.csv` — column definitions
- `bmf_{YYYY_MM}_quality_report.json` — machine-readable quality metrics

<br><hr><br>

## Geography Crosswalks

Optional join layers for the **geocoded Master BMF** that turn a messy address into clean geography: attach a canonical county FIPS code to each organization, then roll that county up to its metro/micropolitan area. Reach for these when you need a collision-proof geographic key — Baltimore **city** (`24510`) is not Baltimore **County** (`24005`) — or want to aggregate nonprofits to metro areas.

These rows **link the canonical S3 copies** — the same files the pipeline consumes — so the catalog never drifts from the machine contract. The CSV is for humans and spreadsheets; the parquet is the machine-typed copy; each `_manifest.json` records the row count, sha256, and geography vintage.

### How the crosswalks chain together

These are **derived, multi-hop** artifacts: you don't start from `county_fips`, you arrive at it. The geocoded BMF gives you a dirty county label and a lat/lon; the crosswalks walk that to a metro area:

1. **County label → county FIPS.** Join the **county FIPS** crosswalk on `(geo_state_abbr, geo_county)` — the crosswalk side of the county key is `geo_county_raw` — to get a canonical 5-char `geo_county_fips`.
2. **Connecticut detour (by coordinate).** CT replaced its counties with planning regions in 2022, so CT's old county labels are deferred in step 1 (`resolution = deferred_ct_planning_region`). Resolve them instead from the org's coordinate: round lat/lon to 2 decimals and join the **CT planning-region** crosswalk on `(lat2, lon2)` to get the planning-region FIPS (`09110`–`09190`).
3. **County FIPS → metro area.** Chain the **CBSA** crosswalk on `county_fips` to attach the metro/micropolitan area (and Combined Statistical Area, where applicable). The CBSA crosswalk already keys CT on the new planning-region FIPS, so step 2 lands cleanly here.

All three share **one geography vintage** (Census TIGER 2023 / OMB 2023) and are designed to be chained — present them together, not in isolation. FIPS/CBSA/CSA codes are **strings**; preserve leading zeros. Unresolved/ambiguous county labels and rural counties carry `NA` codes **by design**, not as missing data.

| Crosswalk | Role in the chain | Geography vintage | Rows | Download | Machine copy | Manifest |
|:---|:---|:---|:---|:---|:---|:---|
| **County FIPS** | label → FIPS (hop 1) | Census TIGER **2023** | 3,635 (3,615 resolved, 99.4%) | <a href='https://nccsdata.s3.us-east-1.amazonaws.com/crosswalks/county-fips/county_fips_crosswalk.csv'>CSV</a> | <a href='https://nccsdata.s3.us-east-1.amazonaws.com/crosswalks/county-fips/county_fips_crosswalk.parquet'>Parquet</a> | <a href='https://nccsdata.s3.us-east-1.amazonaws.com/crosswalks/county-fips/_manifest.json'>manifest.json</a> |
| **CT planning-region** | CT by coordinate (hop 2) | Census TIGER **2023** | 14,271 (coordinate grid) | <a href='https://nccsdata.s3.us-east-1.amazonaws.com/crosswalks/ct-planning-region/ct_planning_region_crosswalk.csv'>CSV</a> | <a href='https://nccsdata.s3.us-east-1.amazonaws.com/crosswalks/ct-planning-region/ct_planning_region_crosswalk.parquet'>Parquet</a> | <a href='https://nccsdata.s3.us-east-1.amazonaws.com/crosswalks/ct-planning-region/_manifest.json'>manifest.json</a> |
| **CBSA** (metro/micro) | FIPS → metro area (hop 3) | OMB **2023** delineations | 3,228 (1,915 in a CBSA, 1,313 rural) | <a href='https://nccsdata.s3.us-east-1.amazonaws.com/crosswalks/cbsa/cbsa_crosswalk.csv'>CSV</a> | <a href='https://nccsdata.s3.us-east-1.amazonaws.com/crosswalks/cbsa/cbsa_crosswalk.parquet'>Parquet</a> | <a href='https://nccsdata.s3.us-east-1.amazonaws.com/crosswalks/cbsa/_manifest.json'>manifest.json</a> |

> **Geography vintage matters.** County boundaries shift — Connecticut replaced its counties with planning regions in 2022 — so a stale vintage is a silent trap. All three crosswalks are built against the **2023** vintage (Census TIGER 2023 boundaries; OMB 2023 CBSA delineations) and carry the year in-column (`tiger_year` / `delineation_year`). Filter against the boundary year you intend, and don't mix vintages across the chain.

### County FIPS crosswalk

Maps the geocoder's dirty county label (`geo_county`) to the canonical Census county FIPS GEOID + `NAMELSAD`, so you can canonicalize names and filter by a collision-proof key. **3,635 rows; 3,615 resolved (99.4%)** — the remainder are 8 ambiguous city-vs-county labels, 8 CT labels deferred to the planning-region crosswalk, and 4 unresolved.

| Column | Description |
|:---|:---|
| `geo_state_abbr` | 2-letter state postal code (join key) |
| `geo_county_raw` | The geocoder's raw, dirty county label (join key) |
| `geo_county_fips` | Canonical 5-char county FIPS GEOID (string — keep leading zeros) |
| `state_fips` | 2-char state FIPS (string) |
| `geo_county_canonical` | Canonical county name (`NAMELSAD`) |
| `resolution` | How the raw label was matched (`resolved`, `ambiguous`, `unresolved`, `deferred_ct_planning_region`) |
| `tiger_year` | Census TIGER boundary vintage (2023) |

### CT planning-region crosswalk

Connecticut's companion to the county crosswalk. Because CT's old county labels can't be trusted post-2022, this resolves CT geography from the **organization's coordinate** instead of its label. It's a **lookup grid** — one row per ~1 km cell (lat/lon rounded to 2 decimals) mapped to one of CT's **9 planning regions** (FIPS `09110`–`09190`). **14,271 rows.**

| Column | Description |
|:---|:---|
| `geo_state_abbr` | Always `CT` |
| `lat2`, `lon2` | Latitude/longitude rounded to 2 decimals (the join key — round the org's coordinate the same way) |
| `geo_county_fips` | CT planning-region FIPS (`09110`–`09190`; joins to the CBSA crosswalk's `county_fips`) |
| `state_fips` | `09` |
| `geo_county_canonical` | Planning-region name |
| `area_share` | Fraction of the cell falling in this planning region (1 when the cell sits fully inside one region) |
| `straddle` | `TRUE` where a cell straddles a region boundary (529 cells); pick the row with the largest `area_share` |
| `tiger_year` | Census TIGER boundary vintage (2023) |

### CBSA crosswalk

The metro/micropolitan rollup. Maps each county FIPS to its OMB Core-Based Statistical Area, so you can aggregate nonprofits to metro areas. **3,228 rows** (one per county GEOID): **1,915 counties fall in a CBSA** — 1,252 metropolitan, 663 micropolitan, across 935 distinct CBSAs (184 of which are grouped into Combined Statistical Areas) — and **1,313 rural counties** carry `NA`.

| Column | Description |
|:---|:---|
| `county_fips` | 5-char county FIPS (string; joins to the county crosswalk's `geo_county_fips`) |
| `county_name` | County name |
| `cbsa_code` | CBSA code (string — keep leading zeros) |
| `cbsa_title` | CBSA name |
| `cbsa_type` | `Metropolitan Statistical Area` or `Micropolitan Statistical Area` |
| `central_outlying` | Whether the county is central or outlying within the CBSA |
| `csa_code` | Combined Statistical Area code (string), if any |
| `csa_title` | Combined Statistical Area name, if any |
| `delineation_year` | OMB delineation vintage (2023) |

**Caveats**

- **FIPS, CBSA, and CSA codes are strings.** Opening a CSV directly in Excel may strip leading zeros on display — read these columns as text.
- **`NA` is a value, not a gap.** **Rural counties** (those in no CBSA) carry `NA` CBSA/CSA codes by design; that's the correct value. Likewise, ambiguous city-vs-county labels (a bare "Baltimore" or "Roanoke") are intentionally left unresolved upstream in the county crosswalk.

### Access via S3

```
s3://nccsdata/crosswalks/county-fips/         # county FIPS crosswalk (.csv + .parquet + _manifest.json)
s3://nccsdata/crosswalks/ct-planning-region/  # CT planning-region crosswalk (.csv + .parquet + _manifest.json)
s3://nccsdata/crosswalks/cbsa/                # CBSA crosswalk (.csv + .parquet + _manifest.json)
```

Each prefix is flat — the vintage is carried in-column and in `_manifest.json` (use its `sha256` to detect changes), not in the path.

Full producer documentation: [County FIPS crosswalk](https://urbaninstitute.github.io/nccs-data-bmf/13-county-fips-crosswalk.html) · [CBSA crosswalk](https://urbaninstitute.github.io/nccs-data-bmf/14-cbsa-crosswalk.html) · [CT planning-region crosswalk](https://urbaninstitute.github.io/nccs-data-bmf/15-ct-planning-region-crosswalk.html).

<br><hr><br>

## Raw Archives

<details>
<summary><strong>Show raw IRS and NCCS files</strong> (for reproducibility only — most users want the products above)</summary>

### Raw Monthly BMF

Untouched IRS monthly releases. Use these when you need to replicate analysis built on a specific raw file or verify upstream NCCS transformations.

**URL pattern:**

```
https://nccsdata.s3.us-east-1.amazonaws.com/raw/bmf/{YYYY-MM}-BMF.csv
```

Replace `{YYYY-MM}` with the desired vintage (e.g. `2026-01`). See the [IRS BMF documentation](https://www.irs.gov/charities-non-profits/exempt-organizations-business-master-file-extract-eo-bmf) for the original schema.

### Raw Legacy BMF

Original NCCS 501CX-NONPROFIT-PX vintages. Each row links the unmodified file and its profile (data dictionary). Most users should prefer the **harmonized legacy BMF** above — it has the same schema as the transformed monthly product.

```{r}
raw_legacy <- build_raw_legacy_section(manifest)
if (nrow(raw_legacy) > 0) {
  knitr::kable(
    raw_legacy,
    escape    = FALSE,
    col.names = c("Year", "Month", "File", "Download", "Profile", "Size"),
    format    = "html",
    row.names = FALSE
  ) |>
    kableExtra::kable_styling()
} else {
  cat("_Raw legacy BMF files are not yet listed in the manifest._\n")
}
```

</details>

<br><hr><br>

## Documentation & Support

- [BMF Processing Guide](https://urbaninstitute.github.io/nccs-data-bmf/index.html) — how NCCS cleans and harmonizes BMF data
- [Quality Reports Index](https://urbaninstitute.github.io/nccs-data-bmf/quality-reports/) — every per-month and per-product quality report
- [Source repository](https://github.com/UrbanInstitute/nccs-data-bmf) — pipeline code and issue tracker
- [Contact](https://nccs.urban.org/nccs/contact/) — questions, corrections, data requests

```{r}
#| results: asis
cat(sprintf("\n<p style='font-size: 0.85em; color: #666; margin-top: 2em;'>Last verified: %s</p>\n",
            format(Sys.Date(), "%Y-%m-%d")))
```
