---
title: "BMF Data Catalog"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| echo: false

library( dplyr )
library( knitr )
library( kableExtra )
library( stringr )
library( pander )

aws_df <- read.csv(
  "AWS-NCCSDATA.csv"
)
source("build-catalog-functions.R")
nccs_legacy_base_url <- "https://urbaninstitute.github.io/nccs-legacy/dictionary/core/core_archive_html/"
aws_base_url <- "https://nccsdata.s3.amazonaws.com/"

state_mapping <- c(
  "AK" = "Alaska",
  "AL" = "Alabama",
  "AR" = "Arkansas",
  "AS" = "American Samoa",
  "AZ" = "Arizona",
  "CA" = "California",
  "CO" = "Colorado",
  "CT" = "Connecticut",
  "DC" = "District of Columbia",
  "DE" = "Delaware",
  "FL" = "Florida",
  "GA" = "Georgia",
  "GU" = "Guam",
  "HI" = "Hawaii",
  "IA" = "Iowa",
  "ID" = "Idaho",
  "IL" = "Illinois",
  "IN" = "Indiana",
  "KS" = "Kansas",
  "KY" = "Kentucky",
  "LA" = "Louisiana",
  "MA" = "Massachusetts",
  "MD" = "Maryland",
  "ME" = "Maine",
  "MI" = "Michigan",
  "MN" = "Minnesota",
  "MO" = "Missouri",
  "MP" = "Northern Mariana Islands",
  "MS" = "Mississippi",
  "MT" = "Montana",
  "NC" = "North Carolina",
  "ND" = "North Dakota",
  "NE" = "Nebraska",
  "NH" = "New Hampshire",
  "NJ" = "New Jersey",
  "NM" = "New Mexico",
  "NV" = "Nevada",
  "NY" = "New York",
  "OH" = "Ohio",
  "OK" = "Oklahoma",
  "OR" = "Oregon",
  "PA" = "Pennsylvania",
  "PR" = "Puerto Rico",
  "RI" = "Rhode Island",
  "SC" = "South Carolina",
  "SD" = "South Dakota",
  "TN" = "Tennessee",
  "TX" = "Texas",
  "UT" = "Utah",
  "Un" = "Unmapped",
  "VA" = "Virginia",
  "VI" = "U.S. Virgin Islands",
  "VT" = "Vermont",
  "WA" = "Washington",
  "WI" = "Wisconsin",
  "WV" = "West Virginia",
  "WY" = "Wyoming"
)
```

The IRS Business Master File (BMF) is the authoritative registry of tax-exempt organizations in the United States. NCCS provides BMF data in several formats to support different research needs.

For background on what the BMF contains, its scope limitations, and guidance on choosing the right dataset, see the [BMF overview page](../datasets/bmf.html).

This catalog is organized by dataset type. Each section includes download links, documentation, and guidance on when to use that dataset.

## Transformed BMF

The Transformed BMF applies standardized cleaning and validation to monthly IRS releases. Each file includes standardized column names, quality flags identifying potential data issues, and documentation of all transformations applied.

**Use when:** You need current BMF data with consistent formatting and documented quality checks.

**Documentation:** [BMF Processing Guide](https://urbaninstitute.github.io/nccs-data-bmf/index.html) — explains how NCCS cleans and transforms raw IRS data, including field definitions and quality checks | [Source code](https://github.com/UrbanInstitute/nccs-data-bmf)

For background on this dataset, see the [BMF overview page](../datasets/bmf.html#transformed-bmf).

### Recent Releases

| Release | Dataset | Data Dictionary | Quality Report |
|:--------|:-------:|:---------------:|:--------------:|
| 2026-01 | [Download](https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/2026_01/bmf_2026_01_processed.csv) | [Dictionary](https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/2026_01/bmf_2026_01_data_dictionary.csv) | [Report](https://urbaninstitute.github.io/nccs-data-bmf/quality-reports/bmf_2026_01_quality_report.html) |
| 2025-12 | [Download](https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/2025_12/bmf_2025_12_processed.csv) | [Dictionary](https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/2025_12/bmf_2025_12_data_dictionary.csv) | [Report](https://urbaninstitute.github.io/nccs-data-bmf/quality-reports/bmf_2025_12_quality_report.html) |
| 2025-11 | [Download](https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/2025_11/bmf_2025_11_processed.csv) | [Dictionary](https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/2025_11/bmf_2025_11_data_dictionary.csv) | [Report](https://urbaninstitute.github.io/nccs-data-bmf/quality-reports/bmf_2025_11_quality_report.html) |
| 2025-10 | [Download](https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/2025_10/bmf_2025_10_processed.csv) | [Dictionary](https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/2025_10/bmf_2025_10_data_dictionary.csv) | [Report](https://urbaninstitute.github.io/nccs-data-bmf/quality-reports/bmf_2025_10_quality_report.html) |
| 2025-09 | [Download](https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/2025_09/bmf_2025_09_processed.csv) | [Dictionary](https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/2025_09/bmf_2025_09_data_dictionary.csv) | [Report](https://urbaninstitute.github.io/nccs-data-bmf/quality-reports/bmf_2025_09_quality_report.html) |

### Older Releases

Transformed BMF files are available monthly from June 2023 through the present. To access older releases, construct URLs using these patterns:

| File Type | URL Pattern |
|:----------|:------------|
| Dataset | `https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/{YYYY_MM}/bmf_{YYYY_MM}_processed.csv` |
| Data Dictionary | `https://nccsdata.s3.us-east-1.amazonaws.com/processed/bmf/{YYYY_MM}/bmf_{YYYY_MM}_data_dictionary.csv` |
| Quality Report | `https://urbaninstitute.github.io/nccs-data-bmf/quality-reports/bmf_{YYYY_MM}_quality_report.html` |

Replace `{YYYY_MM}` with the desired year and month (e.g., `2023_06` for June 2023).

## Unified BMF

The Unified BMF consolidates all historical BMF releases into a single file, with one row per organization that has ever held tax-exempt status. This enables longitudinal analysis without merging multiple annual files.

Features include:

- `ORG_YEAR_FIRST` and `ORG_YEAR_LAST` variables tracking when organizations entered and exited the BMF
- Most recent address geocoded to Census block
- FIPS codes at block, tract, county, and state levels for merging with Census data
- Metropolitan area codes using current CBSA definitions

**Use when:** You need to track organizations over time, build historical sampling frames, or link nonprofit data to Census geographies.

**Coverage:** 1989 through mid-2025 (update pending)

**Related resources:** [NCCS Census Crosswalk](https://nccs.urban.org/nccs/datasets/census/) for aggregating to additional geographic levels

For background on this dataset, see the [BMF overview page](../datasets/bmf.html#unified-bmf).

### Downloads

Files are organized by state/territory. Unmapped records are in the "Unmapped" file.

<a href="https://nccsdata.s3.us-east-1.amazonaws.com/bmf/unified/v1.2/UNIFIED_BMF_V1.2.csv" class="button3">Download Full Unified BMF (v1.2)</a>

<a href="https://nccsdata.s3.amazonaws.com/harmonized/harmonized_data_dictionary.xlsx" class="button3">Download Data Dictionary</a>

```{r}
paths <- aws_df$Key[grepl("harmonized/bmf/unified", aws_df$Key)]
paths <- paths[grepl("_BMF_", paths)]

state_abbrs <- substr(paths, 24, 25)

download_urls <- paste0(aws_base_url, paths)


download_buttons <- make_buttons(urls = download_urls, button_name = "download")
size <- aws_df %>%
  dplyr::filter(Key %in% paths) %>%
  dplyr::mutate(size_mb = paste0(as.character(round(Size / 1000000, 1)), " mb")) %>%
  dplyr::pull("size_mb")

state <- unlist(state_mapping)
catalog <- data.frame(download_buttons, size, state)

rownames(catalog) <- NULL
catalog %>%
  knitr::kable(
    escape = FALSE,
    col.names = c("Download Link", "Size", "State/Territory"),
    format = "html"
  ) %>%
  kableExtra::kable_styling()
```

## Raw BMF Archives

NCCS archives unmodified monthly BMF files as released by the IRS on their [Exempt Organizations Business Master File Extract](https://www.irs.gov/charities-non-profits/exempt-organizations-business-master-file-extract-eo-bmf) page. These files use the IRS's original schema and variable names.

**Use when:** You need to replicate analysis built on raw IRS files from a specific point in time, or require data exactly as the IRS published it.

**Coverage:** June 2023 to present (monthly snapshots)

**Documentation:** [IRS data dictionary](https://www.irs.gov/pub/irs-soi/eo-info.pdf)

For background on this dataset, see the [BMF overview page](../datasets/bmf.html#raw-bmf-archives).

### Downloads

Files are available from the public `nccsdata` S3 bucket using this URL pattern:

`https://nccsdata.s3.us-east-1.amazonaws.com/raw/bmf/{YYYY-MM}-BMF.csv`

**Recent releases:**

| Release | Download |
|:--------|:--------:|
| 2026-01 | [Download](https://nccsdata.s3.us-east-1.amazonaws.com/raw/bmf/2026-01-BMF.csv) |
| 2025-12 | [Download](https://nccsdata.s3.us-east-1.amazonaws.com/raw/bmf/2025-12-BMF.csv) |
| 2025-11 | [Download](https://nccsdata.s3.us-east-1.amazonaws.com/raw/bmf/2025-11-BMF.csv) |
| 2025-10 | [Download](https://nccsdata.s3.us-east-1.amazonaws.com/raw/bmf/2025-10-BMF.csv) |
| 2025-09 | [Download](https://nccsdata.s3.us-east-1.amazonaws.com/raw/bmf/2025-09-BMF.csv) |

For older releases back to June 2023, replace `{YYYY-MM}` in the URL pattern above (e.g., `2023-06-BMF.csv`).

## Legacy BMF

Historical BMF files produced by NCCS between 1989 and 2022. These files retain their original schemas, which vary across years. Data dictionaries are available for most but not all years.

**Use when:** You need to replicate historical NCCS analyses or require BMF data in its original pre-transformation format.

**Coverage:** 1989–2022

**Note:** Column names are not standardized across files. For longitudinal work, consider the [Unified BMF](#unified-bmf) or wait for the forthcoming Harmonized Legacy BMFs (expected Q1 2026).

For background on this dataset, see the [BMF overview page](../datasets/bmf.html#legacy-bmfs).

### Downloads

The files below link to archived BMF files from the [legacy NCCS website](https://urbaninstitute.github.io/nccs-legacy/). Some data dictionaries are incomplete.

```{r}

bmf_legacy_catalog <- construct_catalog(s3_catalog = aws_df,
                             series = "bmf",
                             paths = aws_df$Key)
bmf_legacy_catalog |>
  knitr::kable(escape=FALSE,
               col.names = c(
                 "Download Link",
                 "Data Dictionary",
                 "Size",
                 "Month",
                 "Year"
               ),
               format = "html") |>
   kableExtra::kable_styling()
```

## In Development

The following datasets are currently in progress. Expected timelines are estimates and subject to change.

| Dataset | Description | Expected |
|:--------|:------------|:--------:|
| Harmonized Legacy BMFs | Legacy files (1989–2022) with standardized column names aligned to Transformed BMF schema | Q1 2026 |
| Metadata Tables | Track changes in organization names, addresses, and NTEE codes over time | Q1 2026 |
| Unified BMF v2.0 | Updated Unified BMF with late 2025 data and new architecture | Q1 2026 |

For details on these planned releases, see the [BMF overview page](../datasets/bmf.html#in-development).

## Release Status

| Dataset | Status | Coverage | Documentation |
|:--------|:------:|:---------|:--------------|
| Transformed BMF | Available | June 2023–present | [BMF Processing Guide](https://urbaninstitute.github.io/nccs-data-bmf/index.html) |
| Unified BMF | Available | 1989–mid-2025 | [Data dictionary](https://nccsdata.s3.amazonaws.com/harmonized/harmonized_data_dictionary.xlsx) |
| Raw BMF Archives | Available | June 2023–present | [IRS data dictionary](https://www.irs.gov/pub/irs-soi/eo-info.pdf) |
| Legacy BMFs | Archived | 1989–2022 | Per-file dictionaries above |
| Harmonized Legacy BMFs | Planned (Q1 2026) | 1989–2022 | TBD |
| Metadata Tables | Planned (Q1 2026) | 1989–present | TBD |
| Unified BMF v2.0 | Planned (Q1 2026) | 1989–late 2025 | TBD |

## Contact

Questions or feedback about BMF datasets? Visit our [contact page](https://nccs.urban.org/nccs/contact/).
