---
title: "BMF Data Catalog"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| echo: false

suppressPackageStartupMessages({
  library(dplyr)
  library(knitr)
  library(kableExtra)
  library(stringr)
})

source("R/bmf_helpers.R")

manifest <- read.csv("AWS-BMF.csv", stringsAsFactors = FALSE)

state_mapping <- c(
  # 50 states + DC
  "AL" = "Alabama",          "AK" = "Alaska",
  "AZ" = "Arizona",          "AR" = "Arkansas",
  "CA" = "California",       "CO" = "Colorado",
  "CT" = "Connecticut",      "DE" = "Delaware",
  "DC" = "District of Columbia",
  "FL" = "Florida",          "GA" = "Georgia",
  "HI" = "Hawaii",           "ID" = "Idaho",
  "IL" = "Illinois",         "IN" = "Indiana",
  "IA" = "Iowa",             "KS" = "Kansas",
  "KY" = "Kentucky",         "LA" = "Louisiana",
  "ME" = "Maine",            "MD" = "Maryland",
  "MA" = "Massachusetts",    "MI" = "Michigan",
  "MN" = "Minnesota",        "MS" = "Mississippi",
  "MO" = "Missouri",         "MT" = "Montana",
  "NE" = "Nebraska",         "NV" = "Nevada",
  "NH" = "New Hampshire",    "NJ" = "New Jersey",
  "NM" = "New Mexico",       "NY" = "New York",
  "NC" = "North Carolina",   "ND" = "North Dakota",
  "OH" = "Ohio",             "OK" = "Oklahoma",
  "OR" = "Oregon",           "PA" = "Pennsylvania",
  "RI" = "Rhode Island",     "SC" = "South Carolina",
  "SD" = "South Dakota",     "TN" = "Tennessee",
  "TX" = "Texas",            "UT" = "Utah",
  "VT" = "Vermont",          "VA" = "Virginia",
  "WA" = "Washington",       "WV" = "West Virginia",
  "WI" = "Wisconsin",        "WY" = "Wyoming",
  # U.S. territories
  "AS" = "American Samoa",   "GU" = "Guam",
  "MP" = "Northern Mariana Islands",
  "PR" = "Puerto Rico",      "VI" = "U.S. Virgin Islands",
  # APO/FPO military
  "AA" = "Armed Forces Americas",
  "AE" = "Armed Forces Europe",
  "AP" = "Armed Forces Pacific",
  # Compact of Free Association
  "FM" = "Federated States of Micronesia",
  "MH" = "Marshall Islands",
  "PW" = "Palau",
  # Missing/unknown state
  "ZZ" = "Unknown / unmapped"
)
```

The IRS Business Master File (BMF) is the authoritative registry of tax-exempt organizations in the United States. NCCS publishes BMF data in several forms; this catalog organizes them by what you're trying to do.

For background on what the BMF contains and its scope limitations, see the [BMF overview page](../datasets/bmf.html).

## Which dataset should I use?

| If you need... | Use | Why |
|:---|:---|:---|
| One row per nonprofit, ever (1989–present) | **Master BMF** | Combines every vintage into a single longitudinal frame |
| A specific month's snapshot (2023-06 or later) | **Transformed monthly BMF** | Cleaned and standardized; column schema is stable across months |
| A specific month's snapshot (1989–2022) | **Harmonized legacy BMF** | Same column schema as transformed; aligns historical vintages |
| The unmodified IRS file | **Raw monthly archive** | Untouched IRS releases, for replication of upstream analysis |
| The original NCCS legacy file | **Raw legacy archive** | Original 501CX-NONPROFIT-PX vintages, for reproducibility only |

<br><hr><br>

## Master BMF

One row per EIN spanning 1989 to present, with most-recent values for each organization plus `first_vintage_ym`, `last_vintage_ym`, and `bmf_source` to track provenance. This is the primary product for longitudinal analysis and as a sampling frame.

The **geocoded** variant is the recommended default — it carries the same row schema as the plain file plus lat/lon joined against a deduplicated address table.

```{r}
headline <- build_master_headline_table(manifest)
if (!is.null(headline)) {
  knitr::kable(
    headline,
    escape    = FALSE,
    col.names = c("Variant", "Download", "Dictionary", "Quality Report", "Size"),
    format    = "html",
    row.names = FALSE
  ) |>
    kableExtra::kable_styling()
} else {
  cat("_Master BMF is not yet listed in the manifest._\n")
}
```

### Download by State/Territory

For users who only need a single state, the geocoded Master BMF is also published as per-state marts (50 states + DC, U.S. territories, APO/FPO military codes, Compact of Free Association nations, and a `ZZ` bucket for rows with no resolvable state). Each mart carries lat/lon and shares a schema and dictionary with the geocoded headline file above.

```{r}
master <- build_master_section(manifest, state_mapping)
knitr::kable(
  master,
  escape    = FALSE,
  col.names = c("Download Link", "Size", "State/Territory"),
  format    = "html",
  row.names = FALSE
) |>
  kableExtra::kable_styling()
```

### Access via S3

```
s3://nccsdata/geocoding/bmf-master/merged/   # Geocoded variant (recommended)
s3://nccsdata/master/bmf/                    # Plain Master BMF
s3://nccsdata/master/bmf/state_marts/csv/    # Per-state geocoded marts
```

<br><hr><br>

## Monthly BMF

Cleaned, schema-stable monthly snapshots covering 1989 to the present. The transformed pipeline (2023-06+) and the harmonized legacy pipeline (1989-2022) write to different S3 prefixes but share a column schema, so you can stack them across the era boundary without reshaping.

### Recent Releases

One row per release. The Download column links the CSV (canonical format); a parquet variant is also available for some months at the same S3 prefix.

```{r}
recent <- build_combined_monthly_section(manifest, n_recent = 5)
knitr::kable(
  recent,
  escape    = FALSE,
  col.names = c("Year", "Month", "Pipeline", "Download", "Dictionary", "Quality Report", "Size"),
  format    = "html",
  row.names = FALSE
) |>
  kableExtra::kable_styling()
```

### Browse Older Releases

The complete archive of every monthly release — data, dictionary, and quality report — is below. (There is no separate centralized index; this catalog is the canonical listing.)

<details>
<summary><strong>Show all monthly releases (1989–present)</strong></summary>

```{r}
all_releases <- build_combined_monthly_section(manifest, n_recent = NA)
knitr::kable(
  all_releases,
  escape    = FALSE,
  col.names = c("Year", "Month", "Pipeline", "Download", "Dictionary", "Quality Report", "Size"),
  format    = "html",
  row.names = FALSE
) |>
  kableExtra::kable_styling()
```

</details>

### Access via S3

```
s3://nccsdata/processed/bmf/{YYYY_MM}/         # Transformed (2023-06 -> present)
s3://nccsdata/processed/bmf-legacy/{YYYY_MM}/  # Harmonized legacy (1989-06 -> 2022)
```

Each month directory contains:

- `bmf_{YYYY_MM}_processed.csv` (or `.parquet`) — the data
- `bmf_{YYYY_MM}_data_dictionary.csv` — column definitions
- `bmf_{YYYY_MM}_quality_report.json` — machine-readable quality metrics

<br><hr><br>

## Raw Archives

<details>
<summary><strong>Show raw IRS and NCCS files</strong> (for reproducibility only — most users want the products above)</summary>

### Raw Monthly BMF

Untouched IRS monthly releases. Use these when you need to replicate analysis built on a specific raw file or verify upstream NCCS transformations.

**URL pattern:**

```
https://nccsdata.s3.us-east-1.amazonaws.com/raw/bmf/{YYYY-MM}-BMF.csv
```

Replace `{YYYY-MM}` with the desired vintage (e.g. `2026-01`). See the [IRS BMF documentation](https://www.irs.gov/charities-non-profits/exempt-organizations-business-master-file-extract-eo-bmf) for the original schema.

### Raw Legacy BMF

Original NCCS 501CX-NONPROFIT-PX vintages. Each row links the unmodified file and its profile (data dictionary). Most users should prefer the **harmonized legacy BMF** above — it has the same schema as the transformed monthly product.

```{r}
raw_legacy <- build_raw_legacy_section(manifest)
if (nrow(raw_legacy) > 0) {
  knitr::kable(
    raw_legacy,
    escape    = FALSE,
    col.names = c("Year", "Month", "File", "Download", "Profile", "Size"),
    format    = "html",
    row.names = FALSE
  ) |>
    kableExtra::kable_styling()
} else {
  cat("_Raw legacy BMF files are not yet listed in the manifest._\n")
}
```

</details>

<br><hr><br>

## Documentation & Support

- [BMF Processing Guide](https://urbaninstitute.github.io/nccs-data-bmf/index.html) — how NCCS cleans and harmonizes BMF data
- [Quality Reports Index](https://urbaninstitute.github.io/nccs-data-bmf/quality-reports/) — every per-month and per-product quality report
- [Source repository](https://github.com/UrbanInstitute/nccs-data-bmf) — pipeline code and issue tracker
- [Contact](https://nccs.urban.org/nccs/contact/) — questions, corrections, data requests

```{r}
#| results: asis
cat(sprintf("\n<p style='font-size: 0.85em; color: #666; margin-top: 2em;'>Last verified: %s</p>\n",
            format(Sys.Date(), "%Y-%m-%d")))
```
