---
title: "NCCS 'Core' Data Catalog"
format: html
editor: visual
execute: 
  echo: false
  warning: false
---

```{r}
#| echo: false

library( dplyr )
library( knitr )
library( kableExtra )
library( stringr )
library( pander )

aws_df <- read.csv(
  "AWS-NCCSDATA.csv"
)
source("build-catalog-functions.R")
nccs_legacy_base_url <- "https://urbaninstitute.github.io/nccs-legacy/dictionary/core/core_archive_html/"
aws_base_url <- "https://nccsdata.s3.amazonaws.com/"
```

<br>

<hr>

<br>

Files are organized as follows:

**Tax Exempt Type (tscope)**

-   CHARITIES - 501c3 nonprofit organizations
-   NONPROFITS - All other 501c type organizations besides 501c3
-   PRIVATE FOUNDATIONS - 501c3 private foundations

**IRS 990 Form Scope (fscope)**

Nonprofits can file Form 990, 990EZ, 990PF, or 990N.

Private foundations must file Form 990PF. Small nonprofits can file form 990EZ, which is a shorter version of the full form. Organizations with more than \$250k in revenues must file full Form 990.

Very small nonprofits (revenues less than \$50k annually) can file Form 990N (the postcard form). Since they disclose no finances or organizational activities form 990N filers are not included in the Core Data Series.

Form scope describes which types of filers are included in each data set:

-   **PZ**: 990 + 990EZ filers
-   **PC**: 990 filers only
-   **PF**: 990PF private foundations

Note that scope PZ datasets include a large number of organizations but a smaller number of variables since Form 990EZ has a more limited set of disclosures.

**Year**

Files for form scopes PZ and PC are organized by **tax year**. While PF files are organized by **filing year** See the Research Handbook for details on the differences.

**Data Dictionary**

| Column | Description | Details |
| :---: | :---: | :---: |
| variable_name | Name of variable in CORE data set| |
| variable_description | Variable description taken from concordance | |
| variable_source | Variable Source | HD: Header, PC: Form 990, PZ: Form 990 + 990-EZ |
| form_location | Location on Form 990 | |
| variable_coverage | Tax Years Where Variable Is Present | |
| variable_datatype | Data Type | |

<a href="https://nccsdata.s3.amazonaws.com/harmonized/core/CORE-HRMN_dd.csv" class="button3">Download Dictionary</a>

<br>

<hr>

<br>

#### 501(c)(3) Charities 
### Scope: PZ

```{r}
#| echo: false

paths <- aws_df$Key[grepl("harmonized/core/501c3-pz", aws_df$Key)]
download_urls <- paste0(aws_base_url, paths)
download_buttons <- make_buttons(urls = download_urls, button_name = "download")
size <- aws_df %>%
  dplyr::filter(Key %in% paths) %>%
  dplyr::mutate(size_mb = paste0(as.character(round(Size / 1000000, 1)), " mb")) %>%
  dplyr::pull("size_mb")
year <- get_year(paths)
catalog <- data.frame(download_buttons, size, year)
catalog$np_type <- "501C3-CHARITIES"
catalog$fscope <- "PZ"

catalog %>%
  knitr::kable(
    escape = FALSE,
    col.names = c("Download Link", "Size", "Year", "Nonprofit Type", "Form Scope")
  ) %>%
  kableExtra::kable_styling()
```

<br>

<hr>

<br>

#### 501(c)(3) Charities
### Scope: PC

```{r}
#| echo: false
paths <- aws_df$Key[grepl("harmonized/core/501c3-pc", aws_df$Key)]
download_urls <- paste0(aws_base_url, paths)
download_buttons <- make_buttons(urls = download_urls, button_name = "download")
size <- aws_df %>%
  dplyr::filter(Key %in% paths) %>%
  dplyr::mutate(size_mb = paste0(as.character(round(Size / 1000000, 1)), " mb")) %>%
  dplyr::pull("size_mb")
year <- get_year(paths)
catalog <- data.frame(download_buttons, size, year)
catalog$np_type <- "501C3-CHARITIES"
catalog$fscope <- "PC"

catalog %>%
  knitr::kable(
    escape = FALSE,
    col.names = c("Download Link", "Size", "Year", "Nonprofit Type", "Form Scope")
  ) %>%
  kableExtra::kable_styling()
```

<br>

<hr>

<br>

#### All Other 501(c) Nonprofits
### Scope: PZ

```{r}
#| echo: false
paths <- aws_df$Key[grepl("harmonized/core/501ce-pz", aws_df$Key)]
download_urls <- paste0(aws_base_url, paths)
download_buttons <- make_buttons(urls = download_urls, button_name = "download")
size <- aws_df %>%
  dplyr::filter(Key %in% paths) %>%
  dplyr::mutate(size_mb = paste0(as.character(round(Size / 1000000, 1)), " mb")) %>%
  dplyr::pull("size_mb")
year <- get_year(paths)
catalog <- data.frame(download_buttons, size, year)
catalog$np_type <- "501CE-NONPROFIT"
catalog$fscope <- "PZ"

catalog %>%
  knitr::kable(
    escape = FALSE,
    col.names = c("Download Link", "Size", "Year", "Nonprofit Type", "Form Scope")
  ) %>%
  kableExtra::kable_styling()
```

<br>

<hr>

<br>

#### All Other 501(c) Nonprofits 
### Scope: PC

```{r}
#| echo: false
paths <- aws_df$Key[grepl("harmonized/core/501ce-pc", aws_df$Key)]
download_urls <- paste0(aws_base_url, paths)
download_buttons <- make_buttons(urls = download_urls, button_name = "download")
size <- aws_df %>%
  dplyr::filter(Key %in% paths) %>%
  dplyr::mutate(size_mb = paste0(as.character(round(Size / 1000000, 1)), " mb")) %>%
  dplyr::pull("size_mb")
year <- get_year(paths)
catalog <- data.frame(download_buttons, size, year)
catalog$np_type <- "501CE-NONPROFIT"
catalog$fscope <- "PC"

catalog %>%
  knitr::kable(
    escape = FALSE,
    col.names = c("Download Link", "Size", "Year", "Nonprofit Type", "Form Scope")
  ) %>%
  kableExtra::kable_styling()
```

<br>

<hr>

<br>

#### Private Foundations 
### Scope: PF

```{r}
#| echo: false
catalog <- construct_catalog(s3_catalog = aws_df,
                             series = "core",
                             paths = aws_df$Key,
                             tscope = "PRIVFOUND",
                             fscope = "PF",
                             np_type = "501C3-CHARITIES")
catalog %>%
  knitr::kable( escape=FALSE,
                col.names = c("Download Link", 
                              "Data Dictionary",
                              "Size",
                              "Year",
                              "Nonprofit Type",
                              "Form Scope" ) ) %>%
  kableExtra::kable_styling()
```

<br>

<hr>

<br>

```{=html}
<style>


h1.title {
 font-size: 60px;
 color: #0a4c6a;
}


h4, .h4 {
    font-size: 40px;
    color: #1696d2;
    margin-bottom: 50px;
}



.button {
  background-color: white;
  color:  #12719e;
  border: 2px solid  #12719e;   /* #008CBA; */
  padding: 5px 10px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 14px;
  border-radius: 12px;
  width: 150px;
}
 
.button {
  transition-duration: 0.4s;
}

.button:hover {
  background-color: #1696d2; 
  color: white;
  border: 2px solid #12719e;
}


.button2 {
  background-color: white;
  color: #12719e;
  border: 2px solid #9d9d9d;   /* #008CBA; */
  padding: 5px 10px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 14px;
  border-radius: 12px;
  width: 150px;
}

.button2 {
  transition-duration: 0.4s;
}

.button2:hover {
  background-color: #fff2cf; 
  color: #1696d2;
  font-weight: bold;
  border: 2px solid #1696d2;
}


.center {text-align: center;}

table td {
  vertical-align: middle;
}

blockquote {
    border-left: 4px solid #1696d2;
    margin-left: 10px;
    margin-inline-start: 0px;
    margin-inline-end: 0px;
    padding: 15px;
}

.button3 {
  background-color: #0a4c6a;
  color: white;
  border: 2px solid #0a4c6a;   
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 20px;
  border-radius: 4px;
  width: 200px;
  display: grid;
  margin: auto;
}
</style>
```
