---
title: "Core Data Catalog"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| echo: false

suppressPackageStartupMessages({
  library(dplyr)
  library(knitr)
  library(kableExtra)
  library(stringr)
})

source("R/core_helpers.R")

manifest <- read.csv("AWS-NCCSDATA.csv", stringsAsFactors = FALSE)
```

The NCCS Core Data Series is a panel of nonprofit organizations derived from annual IRS Form 990 filings. This catalog organizes the published files by **processing tier** &mdash; pick the tier that matches your time-span and analytic needs, then drill into a form type within it.

For background on what the Core Series contains and its scope limitations, see the [Core overview page](../datasets/core.html).

::: {.callout-note title="Files are organized by tax year, not filing year"}
Every year label in this catalog refers to the **tax year covered by the return** &mdash; the fiscal period the filing describes &mdash; not the calendar year the IRS received it. A nonprofit's 2022 tax-year return is in the `2022` directory regardless of when it was actually filed (typically 6&ndash;18 months later). Recent tax years are still in-progress as the IRS continues to release returns over time.
:::

## Which tier should I use?

The Core Series is now published in three tiers. Most analysts should default to the **merged panel**; the SOI-current and legacy tiers expose the underlying per-pipeline sources for cases where the merged dedup is too aggressive.

| Tier | Coverage | Forms | When to use |
|:---|:---|:---|:---|
| **Merged panel** &mdash; [jump &darr;](#section-merged) | 1987&ndash;2024 | `990combined`, `990pf` | **Default for most analysts.** One row per `(ein, tax_period)` across the full historical record. SOI takes precedence on overlap years. |
| **SOI-current** &mdash; [jump &darr;](#section-soi) | 2012&ndash;2024 | `990`, `990ez`, `990combined`, `990pf` | When you need the long-form 990 or 990-EZ standalone variable sets, or every filing including amendments (use the `is_amendment` flag). |
| **Legacy** &mdash; [jump &darr;](#section-legacy) | 1987&ndash;2011 | `990combined`, `990pf` | Pre-SOI-era filings, harmonized from the historical NCCS CORE files. Includes all rows from the source files without dedup. |

::: {.callout-warning title="Merged-panel dedup caveat"}
The merged panel is deliberately **one row per `(ein, tax_period)`**, keeping the first occurrence on each side with SOI-current taking precedence over legacy. It is **not** a complete revision history. Analysts who need every original filing plus amendments should use:

- **SOI-current** &mdash; carries the `is_amendment` flag; filter `is_amendment == FALSE` for originals only.
- **Legacy** &mdash; raw harmonized legacy with no dedup, includes every row from the source files.

See [legacy harmonization &amp; merge semantics](https://urbaninstitute.github.io/nccs-data-core/09-legacy-harmonization.html) in the pipeline guide for SOI-precedence rules and the disagreement audit.
:::

<br><hr><br>

## Merged Panel {#section-merged}

Default analyst-facing tier. Column-merge of the legacy and SOI-current panels on `(ein, tax_period)`, one row per filer-period. SOI-current values take precedence; legacy contributes columns where it has them and SOI does not. Includes two tag columns:

- `source_pipeline` &mdash; `"legacy"` or `"soi_current"`, indicates which side a row came from.
- `has_legacy_augment` &mdash; `TRUE` iff legacy contributed columns to a SOI row at merge time (overlap row).

### 990 Combined &mdash; merged {#section-merged-990combined}

Stacked 990 + 990-EZ filers on the 53-column shared variable subset, spanning 1987&ndash;2024.

```{r}
render_year_table(build_core_year_section(manifest, "990combined", tier = "merged"))
```

<details>
<summary><strong>Show all years (merged 990combined)</strong></summary>

```{r}
render_year_table(build_core_year_section(manifest, "990combined", tier = "merged", n_recent = NA))
```

</details>

### 990PF &mdash; merged {#section-merged-990pf}

Private foundations, 1987&ndash;2024. Tax years 1993 and 2017&ndash;2019 are striped: those partitions exist but are mostly empty (see [Known coverage gaps](#known-coverage-gaps) below).

```{r}
render_year_table(build_core_year_section(manifest, "990pf", tier = "merged"))
```

<details>
<summary><strong>Show all years (merged 990pf)</strong></summary>

```{r}
render_year_table(build_core_year_section(manifest, "990pf", tier = "merged", n_recent = NA))
```

</details>

### Access via S3

```
s3://nccsdata/processed_merged/core/{YYYY}/{form_type}/
```

<br><hr><br>

## SOI-Current Panel (2012&ndash;2024) {#section-soi}

The annual SOI-current pipeline: IRS Statistics of Income extracts harmonized into year-by-form-type files. Carries the `is_amendment` flag (`TRUE` iff the `(ein, tax_period)` has appeared in an earlier processing-year extract) and `extract_year` (the processing year the row was sourced from).

### 990 {#section-990}

Full Form 990 filers &mdash; the long-form return required for nonprofits with gross receipts &ge; \$200k or assets &ge; \$500k. Richest variable set in the Core Series (251 columns in 2024).

```{r}
render_year_table(build_core_year_section(manifest, "990", tier = "soi"))
```

<details>
<summary><strong>Show all 990 releases (SOI)</strong></summary>

```{r}
render_year_table(build_core_year_section(manifest, "990", tier = "soi", n_recent = NA))
```

</details>

### 990EZ {#section-990ez}

Short-form filers only. 76 columns in 2024. Most users want **990 Combined** below; this standalone slice is useful when you specifically need only the EZ filers.

```{r}
render_year_table(build_core_year_section(manifest, "990ez", tier = "soi"))
```

<details>
<summary><strong>Show all 990EZ releases (SOI)</strong></summary>

```{r}
render_year_table(build_core_year_section(manifest, "990ez", tier = "soi", n_recent = NA))
```

</details>

### 990 Combined {#section-990combined}

Form 990 + Form 990-EZ filers stacked on the 53-column shared variable subset (plus pipeline-derived universals: `tax_year`, `tax_month`, `is_501c3`, `extract_year`, `is_amendment`).

```{r}
render_year_table(build_core_year_section(manifest, "990combined", tier = "soi"))
```

<details>
<summary><strong>Show all 990 Combined releases (SOI)</strong></summary>

```{r}
render_year_table(build_core_year_section(manifest, "990combined", tier = "soi", n_recent = NA))
```

</details>

### 990PF {#section-990pf}

Private foundations. 187 columns in 2024. Tax years 2017&ndash;2019 are striped: IRS never published 990-PF extracts for those years (see [Known coverage gaps](#known-coverage-gaps)).

```{r}
render_year_table(build_core_year_section(manifest, "990pf", tier = "soi"))
```

<details>
<summary><strong>Show all 990PF releases (SOI)</strong></summary>

```{r}
render_year_table(build_core_year_section(manifest, "990pf", tier = "soi", n_recent = NA))
```

</details>

### Access via S3

```
s3://nccsdata/processed/core/{YYYY}/{form_type}/
```

<br><hr><br>

## Legacy Panel (1987&ndash;2011) {#section-legacy}

Pre-SOI-era filings harmonized from the historical NCCS CORE bucket. No dedup is applied; every row from the source legacy files is preserved. Use this when you need the full pre-2012 record including duplicates.

### 990 Combined &mdash; legacy {#section-legacy-990combined}

```{r}
render_year_table(build_core_year_section(manifest, "990combined", tier = "legacy"))
```

<details>
<summary><strong>Show all years (legacy 990combined)</strong></summary>

```{r}
render_year_table(build_core_year_section(manifest, "990combined", tier = "legacy", n_recent = NA))
```

</details>

### 990PF &mdash; legacy {#section-legacy-990pf}

Tax year 1993 is striped: the source `CORE-1993-501C3-PRIVFOUND-PF.csv` is missing from the legacy NCCS bucket, so the partition contains only ~11k spillover rows.

```{r}
render_year_table(build_core_year_section(manifest, "990pf", tier = "legacy"))
```

<details>
<summary><strong>Show all years (legacy 990pf)</strong></summary>

```{r}
render_year_table(build_core_year_section(manifest, "990pf", tier = "legacy", n_recent = NA))
```

</details>

### Access via S3

```
s3://nccsdata/processed_legacy/core/{YYYY}/{form_type}/
```

<br><hr><br>

## Output schema additions

Columns added by the new pipeline that are worth knowing about &mdash; see the [output schema reference](https://urbaninstitute.github.io/nccs-data-core/08-output-schema.html) and [legacy harmonization guide](https://urbaninstitute.github.io/nccs-data-core/09-legacy-harmonization.html) for the full definitions.

| Column | Tier | Type | What it means |
|:---|:---|:---|:---|
| `is_amendment` | SOI-current | boolean | `TRUE` iff this `(ein, tax_period)` has appeared in an earlier processing-year extract. Filter `== FALSE` for original filings only. |
| `extract_year` | SOI-current | integer | The processing year of the IRS extract this row came from. Useful for amendment provenance. |
| `source_pipeline` | Merged | string | `"legacy"` or `"soi_current"`. Which side of the merge contributed this row. |
| `has_legacy_augment` | Merged | boolean | `TRUE` iff legacy contributed columns to this SOI-current row at merge time (overlap-year augmentation). |
| `is_501c3` | all | boolean | `TRUE` for 501(c)(3) public charities; pair with `subsection_cd` (1&ndash;93) for finer 501(c) subclass breakdowns. |

<br><hr><br>

## Known coverage gaps {#known-coverage-gaps}

These partitions exist in the manifest but reflect IRS-side gaps rather than normal coverage. Rows are striped and flagged with a `&dagger;` in the tables above.

- **SOI 990PF, tax years 2017&ndash;2019** &mdash; IRS never published 990-PF extracts for those tax years. The `processed/core/.../990pf/` partitions contain only late-filer spillover (~665 / 3,348 / 100,359 rows vs. ~100k typical full-year extract). The merged tier inherits this gap.
- **Legacy 990PF, tax year 1993** &mdash; `CORE-1993-501C3-PRIVFOUND-PF.csv` is missing from the legacy NCCS bucket. The `processed_legacy/1993/990pf/` and `processed_merged/1993/990pf/` partitions contain ~11k spillover rows vs. ~40k in 1992 and 1994.

See the [upstream-source quirks](https://urbaninstitute.github.io/nccs-data-core/11-upstream-source-quirks.html) section of the pipeline guide for the full list of IRS-side coverage issues.

<br><hr><br>

## File conventions

Every year/form directory in all three tiers contains the same four artifacts:

```
{TIER_PREFIX}/{YYYY}/{FORM_TYPE}/
    core_{YYYY}_{FORM_TYPE}.csv                # data, CSV
    core_{YYYY}_{FORM_TYPE}.parquet            # data, Parquet (3-5x smaller; fastest for DuckDB / arrow)
    core_{YYYY}_{FORM_TYPE}_dictionary.csv     # variable definitions
    core_{YYYY}_{FORM_TYPE}_dictionary.parquet # variable definitions, Parquet
```

Where `{TIER_PREFIX}` is one of `processed_merged/core`, `processed/core`, or `processed_legacy/core`.

Quality reports are HTML pages served from GitHub Pages, with one URL prefix per tier:

```
https://urbaninstitute.github.io/nccs-data-core/quality-reports/{YYYY}/{form}/core_{YYYY}_{form}_quality.html         # SOI-current
https://urbaninstitute.github.io/nccs-data-core/quality-reports/legacy/{YYYY}/{form}/core_{YYYY}_{form}_quality.html  # Legacy
https://urbaninstitute.github.io/nccs-data-core/quality-reports/merged/{YYYY}/{form}/core_{YYYY}_{form}_quality.html  # Merged
```

This catalog page is the index for those reports &mdash; the **Quality Report** column links the appropriate per-tier URL on every row.

Data dictionaries are per-year because the variable set evolves as the underlying 990 forms change. Always read the dictionary for the year you're using rather than assuming column stability across years.

Filter on **`is_501c3`** (boolean) to subset to 501(c)(3) public charities or their complement. The raw **`subsection_cd`** (IRC EO subsection code, 1&ndash;93) is also present for finer breakdowns.

<br><hr><br>

## Before you analyze

Known upstream-IRS data shapes that look like pipeline bugs but aren't &mdash; documented in the pipeline guide so analysts don't have to rediscover them:

- [No top-coding or winsorization](https://urbaninstitute.github.io/nccs-data-core/08-output-schema.html#no-top-coding-or-winsorization) &mdash; identical maxes across related columns (e.g. `total_contributions` = `total_revenue`) reflect real-world equality for one filer, not capping.
- [~12% of 990-EZ rows fail the balance-sheet identity](https://urbaninstitute.github.io/nccs-data-core/11-upstream-source-quirks.html#filer-supplied-balance-sheet-identity-failures-12-of-990-ez-rows) &mdash; filer-supplied, not a harmonization error.
- [Schedule A `_sec170` / `_sec509` zeros are filer-test-specific](https://urbaninstitute.github.io/nccs-data-core/11-upstream-source-quirks.html#schedule-a-sec170-sec509-blocks-are-filer-test-specific-zeros-are-not-necessarily-missing) &mdash; cross-reference `non_pf_status_reason` before interpreting them as missing data.
- [Filer-supplied negatives on revenue fields](https://urbaninstitute.github.io/nccs-data-core/11-upstream-source-quirks.html#filer-supplied-negatives-on-conceptually-non-negative-fields) &mdash; IRS-supplied as-filed; not amendments.
- [Legacy + SOI merge precedence and disagreement audit](https://urbaninstitute.github.io/nccs-data-core/09-legacy-harmonization.html) &mdash; how the merged panel resolves overlapping `(ein, tax_period)` rows.

Full pipeline guide and per-form quality reports are at the [CORE Pipeline Guide](https://urbaninstitute.github.io/nccs-data-core/).

<br><hr><br>

## Blank IRS Forms Archive {#blank-irs-forms-archive}

Blank IRS Form 990, 990-EZ, 990-PF, their schedules (A&ndash;O and R), and the corresponding instruction booklets are mirrored at `s3://nccsdata/raw/core/forms/`. Each cell below links the blank **Form** and (where IRS publishes one standalone) the **Instr.**uction booklet. Empty cells reflect years IRS did not publish that combination &mdash; not a coverage gap on our side. See the [Forms archive](https://urbaninstitute.github.io/nccs-data-core/02-data-lineage.html) section of the Pipeline Guide for the year-coverage details and refresh SOP.

Direct download URLs follow:

```
https://nccsdata.s3.us-east-1.amazonaws.com/raw/core/forms/<basename>_<YYYY>.pdf
```

A note on schedule instructions: M, N, and parts of I/L have no standalone instruction PDF because IRS folds those instructions into the main `i990_YYYY.pdf` booklet linked under Form 990 below.

### Main Forms

```{r}
forms_main <- build_forms_matrix(manifest, CORE_MAIN_FORMS)
knitr::kable(
  forms_main,
  escape    = FALSE,
  format    = "html",
  row.names = FALSE,
  align     = c("l", rep("c", length(CORE_MAIN_FORMS)))
) |>
  kableExtra::kable_styling(full_width = FALSE)
```

### Schedules

Columns are too narrow for the full labels here, so cells use **F** = blank form, **I** = instruction booklet.

```{r}
#| results: asis
cat("<div style='overflow-x: auto;'>")
forms_sched <- build_forms_matrix(manifest, CORE_SCHEDULES, compact = TRUE)
cat(
  knitr::kable(
    forms_sched,
    escape    = FALSE,
    format    = "html",
    row.names = FALSE,
    align     = c("l", rep("c", length(CORE_SCHEDULES)))
  ) |>
    kableExtra::kable_styling(full_width = FALSE)
)
cat("</div>")
```

<br><hr><br>

## Deprecated: `harmonized/core/` (pre-tier layout)

<details>
<summary><strong>Show legacy <code>harmonized/core/</code> files</strong> (for reproducibility only &mdash; use the [Legacy Panel](#section-legacy) above)</summary>

These files predate the current three-tier layout. They split each year's filers along two axes &mdash; **tax-exempt class** (501c3 charities vs. all other 501c types) and **form scope** (PC = full 990 only, PZ = 990 + 990EZ) &mdash; producing five products. The current pipeline pools these into a single set of by-form-type files with an `is_501c3` column for filtering, so the products below are kept only to support replication of analyses built on the older layout.

**Shared dictionaries** (PC/PZ variants share a single dictionary; PF has its own):

- [PC/PZ data dictionary](https://nccsdata.s3.amazonaws.com/harmonized/core/CORE-HRMN_dd.csv)
- [PF data dictionary](https://nccsdata.s3.amazonaws.com/harmonized/core/501c3-pf/dd/DD-PF-HRMN-V0.csv)

```{r}
render_legacy_table <- function(df) {
  knitr::kable(
    df,
    escape    = FALSE,
    col.names = c("Year", "Download", "Size"),
    format    = "html",
    row.names = FALSE
  ) |>
    kableExtra::kable_styling()
}
```

### 501(c)(3) Charities &mdash; PC (full 990 only)

```{r}
render_legacy_table(build_legacy_core_section(manifest, "501c3-pc"))
```

### 501(c)(3) Charities &mdash; PZ (990 + 990EZ)

```{r}
render_legacy_table(build_legacy_core_section(manifest, "501c3-pz"))
```

### Other 501(c) Nonprofits &mdash; PC (full 990 only)

```{r}
render_legacy_table(build_legacy_core_section(manifest, "501ce-pc"))
```

### Other 501(c) Nonprofits &mdash; PZ (990 + 990EZ)

```{r}
render_legacy_table(build_legacy_core_section(manifest, "501ce-pz"))
```

### Private Foundations &mdash; PF

```{r}
render_legacy_table(build_legacy_core_section(manifest, "501c3-pf"))
```

### Access via S3

```
s3://nccsdata/harmonized/core/{PRODUCT}/marts/
```

where `{PRODUCT}` is one of `501c3-pc`, `501c3-pz`, `501ce-pc`, `501ce-pz`, `501c3-pf`.

</details>

<br><hr><br>

## Documentation &amp; Support

- [Core overview page](../datasets/core.html) &mdash; what's in the Core Series, scope limitations, and merging guidance
- [CORE Pipeline Guide](https://urbaninstitute.github.io/nccs-data-core/) &mdash; root of the pipeline documentation site
- [Architecture overview](https://urbaninstitute.github.io/nccs-data-core/01-architecture.html) &mdash; how the three tiers fit together
- [Output schema reference](https://urbaninstitute.github.io/nccs-data-core/08-output-schema.html) &mdash; column-by-column definitions
- [Legacy harmonization &amp; merge semantics](https://urbaninstitute.github.io/nccs-data-core/09-legacy-harmonization.html) &mdash; SOI precedence, disagreement audit, dedup rules
- [Upstream-source quirks](https://urbaninstitute.github.io/nccs-data-core/11-upstream-source-quirks.html) &mdash; IRS extract format changes and missing-year gaps
- [BMF Data Catalog](catalog-bmf.html) &mdash; Business Master File, for merging organizational attributes (NTEE, 501c type, address) onto Core rows via EIN
- [Census Crosswalk Catalog](catalog-census_crosswalk.html) &mdash; aggregate Core data to tracts, counties, or MSAs
- [Contact](https://nccs.urban.org/nccs/contact/) &mdash; questions, corrections, data requests

```{r}
#| results: asis
cat(sprintf("\n<p style='font-size: 0.85em; color: #666; margin-top: 2em;'>Last verified: %s</p>\n",
            format(Sys.Date(), "%Y-%m-%d")))
```
