---
title: "BMF Data Catalog"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| echo: false

suppressPackageStartupMessages({
  library(dplyr)
  library(knitr)
  library(kableExtra)
  library(stringr)
})

source("R/bmf_helpers.R")

manifest <- read.csv("AWS-BMF.csv", stringsAsFactors = FALSE)

state_mapping <- c(
  "AL" = "Alabama",          "AK" = "Alaska",
  "AZ" = "Arizona",          "AR" = "Arkansas",
  "CA" = "California",       "CO" = "Colorado",
  "CT" = "Connecticut",      "DE" = "Delaware",
  "DC" = "District of Columbia",
  "FL" = "Florida",          "GA" = "Georgia",
  "HI" = "Hawaii",           "ID" = "Idaho",
  "IL" = "Illinois",         "IN" = "Indiana",
  "IA" = "Iowa",             "KS" = "Kansas",
  "KY" = "Kentucky",         "LA" = "Louisiana",
  "ME" = "Maine",            "MD" = "Maryland",
  "MA" = "Massachusetts",    "MI" = "Michigan",
  "MN" = "Minnesota",        "MS" = "Mississippi",
  "MO" = "Missouri",         "MT" = "Montana",
  "NE" = "Nebraska",         "NV" = "Nevada",
  "NH" = "New Hampshire",    "NJ" = "New Jersey",
  "NM" = "New Mexico",       "NY" = "New York",
  "NC" = "North Carolina",   "ND" = "North Dakota",
  "OH" = "Ohio",             "OK" = "Oklahoma",
  "OR" = "Oregon",           "PA" = "Pennsylvania",
  "RI" = "Rhode Island",     "SC" = "South Carolina",
  "SD" = "South Dakota",     "TN" = "Tennessee",
  "TX" = "Texas",            "UT" = "Utah",
  "VT" = "Vermont",          "VA" = "Virginia",
  "WA" = "Washington",       "WV" = "West Virginia",
  "WI" = "Wisconsin",        "WY" = "Wyoming",
  "AS" = "American Samoa",   "GU" = "Guam",
  "MP" = "Northern Mariana Islands",
  "PR" = "Puerto Rico",      "VI" = "U.S. Virgin Islands",
  "Un" = "Unmapped"
)
```

The IRS Business Master File (BMF) is the authoritative registry of tax-exempt organizations in the United States. NCCS provides BMF data in several formats to support different research needs.

For background on what the BMF contains, its scope limitations, and guidance on choosing the right dataset, see the [BMF overview page](../datasets/bmf.html).

This catalog is organized by dataset type. Each section includes download links, documentation, and guidance on when to use that dataset.

## Transformed BMF

The BMF catalog has three sections. The **Master BMF** is the unified state-level dataset (formerly "Unified BMF"). **Processed BMF** and **Legacy BMF** provide monthly snapshots, each linked to a quality report.

<a href="https://nccsdata.s3.amazonaws.com/harmonized/harmonized_data_dictionary.xlsx" class="button3">Download Data Dictionary</a>

<br>

<hr>

<br>

#### Master BMF by State/Territory

```{r}
master <- build_master_section(manifest, state_mapping)

knitr::kable(
  master,
  escape    = FALSE,
  col.names = c("Download Link", "Size", "State/Territory"),
  format    = "html",
  row.names = FALSE
) |>
  kableExtra::kable_styling()
```

<br>

<hr>

<br>

## Processed BMF (Monthly)

Monthly processed snapshots from `processed/bmf/`. Each row links the dataset and its quality report.

```{r}
processed <- build_monthly_section(manifest, "processed")

knitr::kable(
  processed,
  escape    = FALSE,
  col.names = c("Year", "Month", "Download", "Quality Report", "Size"),
  format    = "html",
  row.names = FALSE
) |>
  kableExtra::kable_styling()
```

<br>

<hr>

<br>

## Legacy BMF (Monthly)

Monthly legacy snapshots from `processed/bmf-legacy/`. NCCS has since shifted to compiling a Master BMF that aggregates the various BMF fields.

```{r}
legacy <- build_monthly_section(manifest, "legacy")

knitr::kable(
  legacy,
  escape    = FALSE,
  col.names = c("Year", "Month", "Download", "Quality Report", "Size"),
  format    = "html",
  row.names = FALSE
) |>
  kableExtra::kable_styling()
```

## In Development

The following datasets are currently in progress. Expected timelines are estimates and subject to change.

| Dataset | Description | Expected |
|:--------|:------------|:--------:|
| Harmonized Legacy BMFs | Legacy files (1989–2022) with standardized column names aligned to Transformed BMF schema | Q1 2026 |
| Metadata Tables | Track changes in organization names, addresses, and NTEE codes over time | Q1 2026 |
| Unified BMF v2.0 | Updated Unified BMF with late 2025 data and new architecture | Q1 2026 |

For details on these planned releases, see the [BMF overview page](../datasets/bmf.html#in-development).

## Release Status

| Dataset | Status | Coverage | Documentation |
|:--------|:------:|:---------|:--------------|
| Transformed BMF | Available | June 2023–present | [BMF Processing Guide](https://urbaninstitute.github.io/nccs-data-bmf/index.html) |
| Unified BMF | Available | 1989–mid-2025 | [Data dictionary](https://nccsdata.s3.amazonaws.com/harmonized/harmonized_data_dictionary.xlsx) |
| Raw BMF Archives | Available | June 2023–present | [IRS data dictionary](https://www.irs.gov/pub/irs-soi/eo-info.pdf) |
| Legacy BMFs | Archived | 1989–2022 | Per-file dictionaries above |
| Harmonized Legacy BMFs | Planned (Q1 2026) | 1989–2022 | TBD |
| Metadata Tables | Planned (Q1 2026) | 1989–present | TBD |
| Unified BMF v2.0 | Planned (Q1 2026) | 1989–late 2025 | TBD |

## Contact

Questions or feedback about BMF datasets? Visit our [contact page](https://nccs.urban.org/nccs/contact/).
